---
title: "T7 warmx soil temperature Plots"
author: "Adrian Noecker"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

COLLABORATORS: Kara Dobson, Phoebe Zarnetske, Mark Hammond, Moriah Young
DATA INPUT:  csv from the shared Google drive  
DATA OUTPUT: Code and Rmd are in the scripts folder in Github  
PROJECT: REX

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

# Load packages
library(tidyverse)
library(ggpubr)
library(RColorBrewer)
library(plotrix)

# Set working directory from .Renviron
dir <- Sys.getenv("DATA_DIR")
list.files(dir)
setwd("G:/Shared drives/KBS_LTER_REX/data/sensors/OTC Footprints/L1") # MLY - you don't need this anymore since you have the .Renviron working now

# Read in data
soiltemp <- read.csv("T7_warmx_soil_sensors_L1.csv") # MLY - read.csv is also a little different now that you have .Renviron working (see blow code)
soiltemp <- read.csv(file.path(dir, "sensors/OTC Footprints/L1/T7_warmx_soil_sensors_L1.csv")) 

# re-making the data column a date
soiltemp$Date_Time <- as.POSIXct(soiltemp$sample_datetime, format = "%Y-%m-%d %H:%M:%S")
str(soiltemp)

# Specifying the order that we want the treatments to be listed in the figures on the x-axis
level_order <- c('irrigated_control', 'ambient', 'drought', 'warmed','warmed_drought')

```

```{r}
# create new dataframe with only data from 7 AM - 7 PM 
soil_season <- soiltemp
soil_season$date <- paste0(soil_season$month,"",soil_season$day)
soil_season$date <- as.numeric(soil_season$date)
# this is mainly what we'll work with for figures (could always adjust the parameters as needed)
soil_season$month <- format(soil_season$Date_Time,format="%m")
soil_season$year <- format(soil_season$Date_Time,format="%Y")
soil_season$hour <- format(soil_season$Date_Time, format="%H")
soil_season$day <- format(soil_season$Date_Time, format="%d")
soil_season_22 <- soil_season %>%
  filter(year == 2022) %>%
  filter(hour > "06") %>%
  filter(hour < "20")
```

```{r}
# take average per month, per rep, per treatment
soil_sam <- soil_season_22 %>%
  group_by(month, Subplot_Descriptions) %>%
  summarize(avg_temp = mean(temperature, na.rm = TRUE),
            se_temp = std.error(temperature, na.rm = TRUE))
           
```

```{r}
# plot - monthly
#png("rex_soiltemp_22.png", units="in", width=7, height=4, res=300)
ggplot(soil_sam, aes(x = month, y = avg_temp, group = Subplot_Descriptions, color = Subplot_Descriptions)) +
  geom_errorbar(aes(ymin=avg_temp-se_temp, ymax=avg_temp+se_temp),width=0.1,color="black",linetype="solid", position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping error bars - the number is how much you want it to dodge by
  geom_line(linewidth = 1) +
  geom_point(linewidth = 2, position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  scale_color_manual(name="Treatment",
                     values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E", "#6B8E23"),
                     labels=c("Ambient","Drought","Warmed", "Warmed Drought", "Irrigated Control")) +
  labs(y="Soil Temperature (°C)", x="Month") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15),
        legend.title = element_text(size=17),
        legend.text = element_text(size=15))
#dev.off()


```

Making plot that only includes growing season
```{r}
# making new data frame that only includes data during day (7am-7pm) from April to August
soil_season$month <- format(soil_season$Date_Time,format="%m")
soil_season$year <- format(soil_season$Date_Time,format="%Y")
soil_season$hour <- format(soil_season$Date_Time, format="%H")
soil_season$day <- format(soil_season$Date_Time, format="%d")
soil_grow_season_22 <- soil_season %>%
  filter(year == 2022) %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  filter(month > "03") %>% # MLY - changed this to March because it wasn't including April since it was selecting everything greater than April with filter(month > "04")
  filter(month < "09")
```

```{r}
# take average per month, per rep, per treatment
soil_temp_grow <- soil_grow_season_22 %>%
  group_by(month, Subplot_Descriptions) %>%
  summarize(avg_temp = mean(temperature, na.rm = TRUE),
            se_temp = std.error(temperature, na.rm = TRUE))
```

Average monthly soil temperature plot from April - August
```{r}
# plot - monthly
#png("rex_soiltemp_22.png", units="in", width=7, height=4, res=300)
ggplot(soil_temp_grow, aes(x = month, y = avg_temp, group = Subplot_Descriptions, color = Subplot_Descriptions)) +
  geom_errorbar(aes(ymin=avg_temp-se_temp, ymax=avg_temp+se_temp), width=0.1, color="black", linetype="solid") +
  geom_line(linewidth = 1) +
  geom_point(linewidth = 2) +
  scale_color_manual(name="Treatment",
                     values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E", "#6B8E23"),
                     labels=c("Ambient","Drought","Warmed", "Warmed Drought", "Irrigated Control")) +
  labs(y="Soil Temperature (°C)", x="Month") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15),
        legend.title = element_text(size=17),
        legend.text = element_text(size=15))
#dev.off()
```

Adrian - 
        Great job on these graphs! We're seeing something interesting here. Even with using the "dodge" code, there's still a lot of overlap with the lines, but that's ok, that's important bc it's telling us there they are pretty much the same. 
        You don't have to do this in the script, but start thinking about why we are seeing what we are in these graphs. Why might drought have the highest soil temperature? It's interesting that our two subplot manipulations with the OTCs have the lowest avg soil temperature in July and August. Why might this be? Is it related to plant growth/abundance?

Here are the dates of the post and peak droughts from last year if you want to make graphs from those weeks too. You could make one averaging the soil temps from each of these weeks and try your hand at creating code to combine the two graphs so that they are side by side.
Peak Drought happened during the week of: 8/1/2022
Post Drought happened during the week of: 8/8/2022
```{r}

```

