---
title: "T7 warmx air temperature plots"
author: "Adrian Noecker, Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

COLLABORATORS: Kara Dobson, Phoebe Zarnetske, Mark Hammond
DATA INPUT:  csv from the shared Google drive  
DATA OUTPUT: Code and Rmd are in the scripts folder in Github  
PROJECT: REX

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

# Load packages
library(tidyverse)
library(dplyr)
library(ggpubr)
library(RColorBrewer)
library(plotrix)

# Set working directory from .Renviron
dir <- Sys.getenv("DATA_DIR")
list.files(dir)

#read in data
airtemp <- read.csv(file.path(dir, "sensors/OTC Footprints/L1/T7_warmx_HOBO_L1.csv")) 

# re-making the data column a date
airtemp$Date_Time <- as.POSIXct(airtemp$Date_Time, format = "%Y-%m-%d %H:%M:%S")
str(airtemp)

```

```{r}
# create new dataframe with only data from 7 AM - 7 PM 
day_airtemp <- airtemp
day_airtemp$date <- paste0(day_airtemp$month,"",day_airtemp$day)
day_airtemp$date <- as.numeric(day_airtemp$date)
# this is mainly what we'll work with for figures (could always adjust the parameters as needed)
day_airtemp$month <- format(day_airtemp$Date_Time,format="%m")
day_airtemp$year <- format(day_airtemp$Date_Time,format="%Y")
day_airtemp$hour <- format(day_airtemp$Date_Time, format="%H")
day_airtemp$day <- format(day_airtemp$Date_Time, format="%d")
day_airtemp_22 <- day_airtemp %>%
  filter(year == 2022) %>%
  filter(hour > "06") %>%
  filter(hour < "20")
```

```{r}
# take average per month by treatment
avg_airtemp_22 <- day_airtemp_22 %>%
  group_by(month, Treatment) %>%
  summarise(avg_temp = mean(Temperature_C, na.rm = TRUE),
            se_temp = std.error(Temperature_C, na.rm = TRUE))
           
```

```{r}
# plot - monthly
png("rex_airtemp_22.png", units="in", width=8, height=6, res=300)
ggplot(avg_airtemp_22, aes(x = month, y = avg_temp, group = Treatment, color = Treatment)) +
  geom_errorbar(aes(ymin=avg_temp-se_temp, ymax=avg_temp+se_temp),width=0.1,color="black",linetype="solid") +
  geom_line(linewidth = 1) +
  scale_color_manual(name="Treatment",
                     values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E"),
                     labels=c("Ambient","Drought","Warmed", "Warmed Drought")) +
  labs(y="Air Temperature (°C)", x="Month") +
  theme_bw() +
  theme(axis.title = element_text(size=15),
        axis.text = element_text(size=15))
dev.off()
```

Making plot that only includes growing season
```{r}
# making new data frame that only includes data during day (7am-7pm) from April to August
day_airtemp$month <- format(day_airtemp$Date_Time,format="%m")
day_airtemp$year <- format(day_airtemp$Date_Time,format="%Y")
day_airtemp$hour <- format(day_airtemp$Date_Time, format="%H")
day_airtemp$day <- format(day_airtemp$Date_Time, format="%d")
day_airtemp_22 <- day_airtemp %>%
  filter(year == 2022) %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  filter(month > "03") %>% 
  filter(month < "09")
```

```{r}
# take average per month, per rep, per treatment
air_temp_grow <- day_airtemp_22 %>%
  dplyr::group_by(month, Treatment) %>%
  summarize(avg_temp = mean(Temperature_C, na.rm = TRUE),
            se_temp = std.error(Temperature_C, na.rm = TRUE))
```

Average monthly soil temperature plot from April - August
```{r}
# plot - monthly
png("rex_airtemp_growing_season_22.png", units="in", width=6, height=6, res=300)
air_growing_season <- 
        ggplot(air_temp_grow, aes(x = month, y = avg_temp, group = Treatment, color = Treatment)) +
        geom_line(linewidth = 1) +
        geom_errorbar(aes(ymin=avg_temp-se_temp, ymax=avg_temp+se_temp), width=0.1, color="black", linetype="solid") +
        scale_color_manual(name="Treatment",
                     values = c("#a6bddb", "#687689", "#fb6a4a", "#9D422E"),
                     labels=c("Ambient","Drought","Warmed", "Warmed Drought")) +
        labs(y="Air Temperature (°C)", x="Month") +
        theme_bw() +
        theme(axis.title = element_text(size=17),
              axis.text = element_text(size=15))
dev.off()
air_growing_season
```

######################
Adding soil temp below but separate script already exists
```{r}
# Read in data
soiltemp <- read.csv(file.path(dir, "sensors/OTC Footprints/L1/T7_warmx_soil_sensors_L1.csv")) 

# re-making the data column a date
soiltemp$Date_Time <- as.POSIXct(soiltemp$sample_datetime, format = "%Y-%m-%d %H:%M:%S")
str(soiltemp)

# create new dataframe with only data from 7 AM - 7 PM 
soil_season <- soiltemp
soil_season$date <- paste0(soil_season$month,"",soil_season$day)
soil_season$date <- as.numeric(soil_season$date)
# this is mainly what we'll work with for figures (could always adjust the parameters as needed)
soil_season$month <- format(soil_season$Date_Time,format="%m")
soil_season$year <- format(soil_season$Date_Time,format="%Y")
soil_season$hour <- format(soil_season$Date_Time, format="%H")
soil_season$day <- format(soil_season$Date_Time, format="%d")
soil_grow_season_22 <- soil_season %>%
  filter(year == 2022) %>%
  filter(hour > "06") %>%
  filter(hour < "20") %>%
  filter(month > "03") %>% 
  filter(month < "09")

# take average per month, per rep, per treatment
soil_temp_grow <- soil_grow_season_22 %>%
  group_by(month, Subplot_Descriptions) %>%
  summarize(avg_temp = mean(temperature, na.rm = TRUE),
            se_temp = std.error(temperature, na.rm = TRUE))

soil_growing_season <- 
        ggplot(soil_temp_grow, aes(x = month, y = avg_temp, group = Subplot_Descriptions, color = Subplot_Descriptions)) +
        geom_line(linewidth = 1) +
        geom_errorbar(aes(ymin=avg_temp-se_temp, ymax=avg_temp+se_temp), width=0.1, color="black", linetype="solid") +
        scale_color_manual(name="Treatment",
                     values = c("#00008B","#a6bddb", "#687689", "#fb6a4a", "#9D422E"),
                     labels=c("Irrigated Control", "Ambient","Drought","Warmed", "Warmed Drought")) +
        labs(y="Soil Temperature (°C)", x= "Month") +
        theme_bw() +
        theme(axis.title = element_text(size=17),
              axis.text = element_text(size=15))
soil_growing_season
```

combining air and soil temp growing season graphs in one
```{r}
figure <- ggarrange(air_growing_season, soil_growing_season,
                    ncol = 2, nrow = 1)

figure

png("rex_air_soil_temp_2022.png", units="in", width=8, height=5, res=300)
ggpubr::ggarrange(air_growing_season, soil_growing_season,
                                nrow = 1, ncol = 2, common.legend = T, legend = "right", label.x = "Month",
                                align = "v")
dev.off()
```


# for REX methods paper
# First, plotting all the HOBO air temp data
```{r}
# create date, month, year, hour, and day columns
airtemp <- na.omit(airtemp)
airtemp2 <- airtemp
airtemp2$date <- format(airtemp2$Date_Time,format="%Y-%m-%d")
airtemp2$month <- format(airtemp2$Date_Time,format="%m")
airtemp2$year <- format(airtemp2$Date_Time,format="%Y")
airtemp2$month_year <- format(airtemp2$Date_Time,format="%m-%Y")
airtemp2$hour <- format(airtemp2$Date_Time, format="%H")
airtemp2$day <- format(airtemp2$Date_Time, format="%d")

# take average per month by treatment
avg_airtemp2 <- airtemp2 %>%
  group_by(month, year, month_year, Treatment) %>%
  summarise(avg_temp = mean(Temperature_C, na.rm = TRUE),
            se_temp = std.error(Temperature_C, na.rm = TRUE))
```

```{r}
# plot - monthly
png("rex_airtemp_all_years.png", units="in", width=10, height=6, res=300)
ggplot(avg_airtemp2, aes(x = month, y = avg_temp, group = Treatment, color = Treatment)) +
        geom_errorbar(aes(ymin=avg_temp-se_temp, ymax=avg_temp+se_temp),width=0.1,color="black",linetype="solid") +
        geom_line(linewidth = 1) +
        facet_grid(~year) +
        scale_color_manual(name="Treatment", 
                           values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E"), 
                           labels=c("Ambient","Drought","Warmed", "Warmed Drought")) +
        labs(y="Air Temperature (°C)", x="Month") +
        theme_bw() +
        theme(axis.title = element_text(size=12),
        axis.text = element_text(size=12),
        axis.text.x = element_text(angle=45, vjust=1, hjust=1))
dev.off()
```

# Next, select for the dates where the shelters were on
# Need to double check these dates
2021: July 12th - August 27th
2022: June 27th - August 5th
2023: July 10th - August 18th
```{r}
# 2021
airtemp21_drought <- airtemp
airtemp21_drought$month <- format(airtemp21_drought$Date_Time,format="%m")
airtemp21_drought$year <- format(airtemp21_drought$Date_Time,format="%Y")
airtemp21_drought$hour <- format(airtemp21_drought$Date_Time, format="%H")
airtemp21_drought$day <- format(airtemp21_drought$Date_Time, format="%d")
airtemp21_drought$month_year <- format(airtemp21_drought$Date_Time, format="%m-%Y")
airtemp21_drought$date <- format(airtemp21_drought$Date_Time, format="%Y-%m-%d")
airtemp21_drought1 <- airtemp21_drought %>%
  filter(date >= "2021-07-12" & date <= "2021-08-27")

# 2022
airtemp22_drought <- airtemp
airtemp22_drought$month <- format(airtemp22_drought$Date_Time,format="%m")
airtemp22_drought$year <- format(airtemp22_drought$Date_Time,format="%Y")
airtemp22_drought$hour <- format(airtemp22_drought$Date_Time, format="%H")
airtemp22_drought$day <- format(airtemp22_drought$Date_Time, format="%d")
airtemp22_drought$month_year <- format(airtemp22_drought$Date_Time, format="%m-%Y")
airtemp22_drought$date <- format(airtemp22_drought$Date_Time, format="%Y-%m-%d")
airtemp22_drought1 <- airtemp22_drought %>%
  filter(date >= "2022-06-27" & date <= "2022-08-05")

# 2023
airtemp23_drought <- airtemp
airtemp23_drought$month <- format(airtemp23_drought$Date_Time,format="%m")
airtemp23_drought$year <- format(airtemp23_drought$Date_Time,format="%Y")
airtemp23_drought$hour <- format(airtemp23_drought$Date_Time, format="%H")
airtemp23_drought$day <- format(airtemp23_drought$Date_Time, format="%d")
airtemp23_drought$month_year <- format(airtemp23_drought$Date_Time, format="%m-%Y")
airtemp23_drought$date <- format(airtemp23_drought$Date_Time, format="%Y-%m-%d")
airtemp23_drought1 <- airtemp23_drought %>%
  filter(date >= "2023-07-10" & date <= "2023-08-18")

# merge data sets together
peakdrought <- full_join(airtemp22_drought1, airtemp23_drought1)

# take average per month by treatment
avg_peakdrought <- peakdrought %>%
  group_by(year, Treatment) %>%
  summarise(avg_temp = mean(Temperature_C, na.rm = TRUE),
            se_temp = std.error(Temperature_C, na.rm = TRUE))
```

# work in progress
# might work better as a box plot? Should temps be average over the 6 weeks, or do we want to show by month, date, etc?
# Peak Drought Air Temp Plot
```{r}
png("rex_airtemp_peakdrought_allyears.png", units="in", width=6, height=6, res=300)
ggplot(avg_peakdrought, aes(x = year, y = avg_temp, fill = Treatment)) +
        geom_pointrange(aes(ymin=avg_temp-se_temp, ymax=avg_temp+se_temp), pch = 21,size = 1.2,
                        position=position_dodge(0.5)) +
        #facet_grid(~year) +
        scale_fill_manual(name="Treatment", 
                           values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E"), 
                           labels=c("Ambient","Drought","Warmed", "Warmed Drought")) +
        labs(y="Air Temperature (°C)", x="Year") +
        theme_bw() +
        theme(axis.title = element_text(size=15),
        axis.text = element_text(size=15),
        axis.text.x = element_text(angle=45, vjust=1, hjust=1))
dev.off()
```
