---
title: "REX: T7 Red Clover ANPP Analyses"
author: "Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

COLLABORATORS: Kara Dobson, Phoebe Zarnetske, Mark Hammond
DATA INPUT:  csv from the shared Google drive  
DATA OUTPUT: Code and Rmd are in the scripts folder in Github  
PROJECT: REX  

DESCRIPTION:
- Looking at legume species (primarily red clover aka Trifolium pratense - Trpr)
- Main question: have the treatments (warming & drought) altered the amount of biomass and/or within certain species such as legumes?
- 2019 is baseline data taken by Mark Hammond before any treatments were imposed (collected all biomass within the 1m2 plot)
- 2021 is data from the first year of REX (collected biomass within a 0.20m2 area inside the 1m2 plot)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

# Load packages
library(bbmle)
library(lmerTest)
library(fitdistrplus)
library(sjPlot)
library(car)
library(emmeans)
library(tidyverse)
library(ggpubr)
library(jtools) #summ 
library(rstatix)
library(RColorBrewer)
library(plotrix)

# Set working directory from .Renviron
dir <- Sys.getenv("DATA_DIR")
list.files(dir)

# Read in data
anpp19 <- read.csv(file.path(dir, "T7_ANPP/L1/T7_warmx_ANPP_2019_L1.csv")) # 2019 ANPP data
anpp21 <- read.csv(file.path(dir, "T7_ANPP/L1/T7_warmx_ANPP_2021_L1.csv")) # 2021 ANPP data
meta <- read.csv(file.path(dir, "REX_warmx_metadata.csv"))

# Specifying the order that we want the treatments to be listed in the figures on the x-axis
level_order <- c('irrigated_control', 'ambient', 'drought','warmed','warmed_drought')
```

```{r}
# cleaning 2019 data
names(anpp19)[names(anpp19)=="Dried_plant_biomass_g_per_metersquare"] <- "Biomass" # changing column name
names(anpp19)[names(anpp19)=="Group_Code"] <- "Code" # changing column name
names(anpp19)[names(anpp19)=="Suplot_Location"] <- "Subplot_Location" # changing column name
anpp19 <- anpp19[,-c(2, 5, 11:13)] # delete unnecessary columns

# cleaning 2021 data
anpp21 <- anpp21[,-c(5, 8, 11:19)] # delete unnecessary columns
anpp21$Year <- 2021 # add year column
names(anpp21)[names(anpp21)=="Dried_Plant_Biomass_g"] <- "Biomass" # changing column name
names(anpp21)[names(anpp21)=="Species_Code"] <- "Code" # changing column name
# multiply biomass values by 5 - biomass was only clipped at 0.20m2 of the 1m2 plot
anpp21$biomass_scaled <- anpp21$Biomass*5
anpp21_scaled <- anpp21[,-6] # delete column
names(anpp21_scaled)[9] <- "Biomass" #changing column name

# merge 2019 and 2021 data together
anpp_all <- full_join(anpp19, anpp21_scaled)
```

```{r}
# removing treatments we aren't interested in
anpp21 <- anpp21 %>%
  filter(!(Subplot_Descriptions=='drought_insecticide' |
             Subplot_Descriptions=='insecticide' |
             Subplot_Descriptions=='warmed_drought_insecticide' |
             Subplot_Descriptions=='warmed_insecticide'))

# Take subplot average of biomass
anpp21_avg <- anpp21 %>%
        group_by(Subplot_Descriptions) %>%
        summarise(avg_biomass = mean(Biomass, na.rm = TRUE),
        se = std.error(Biomass, na.rm = TRUE))

png("avg_anpp_2021.png", units="in", width=6, height=6, res=300)
ggplot(anpp21_avg, aes(x = Subplot_Descriptions, y = avg_biomass, fill = Subplot_Descriptions)) +
        geom_jitter(shape=16, position=position_jitterdodge(), alpha = 0.6, aes(colour = Subplot_Descriptions)) +
        geom_bar(stat = "identity") +
        geom_errorbar(aes(ymin = avg_biomass - se, ymax = avg_biomass + se), width = 0.2,
                         position = "identity") +
        scale_fill_manual(values = c("ambient" = "darkblue", "drought" = "gray58", 
        "irrigated_control" = "deepskyblue", "warmed" = "red2", "warmed_drought" = "darkred")) +
        labs(x = "Treatment", y = "2021 Average Biomass (g/0.20m^2)") +
        scale_x_discrete(limits = level_order,
                         labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "irrigated_control" = "Irrigated \n Control",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none")
dev.off

ggplot(anpp21, aes(x = Subplot_Descriptions, y = Biomass, fill = Subplot_Descriptions)) +
        geom_jitter(shape=16, position=position_jitterdodge(), alpha = 0.6, aes(colour = Subplot_Descriptions)) +
        geom_boxplot(color = "black", outlier.shape = NA) +
        scale_fill_manual(values = c("ambient" = "darkblue", "drought" = "gray58", 
        "irrigated_control" = "deepskyblue", "warmed" = "red2", "warmed_drought" = "darkred")) +
        labs(x = "Treatment", y = "2021 Average Biomass (g/0.20m^2)") +
        scale_x_discrete(limits = level_order,
                         labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "irrigated_control" = "Irrigated \n Control",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none")
```

```{r}
# select only legume data from ANPP 2019
legumes <- anpp_all[(anpp_all$Code == "Legume" | anpp_all$Code == "TRFPR"),]

# Take treatment average total biomass
legumes_avg <- legumes %>%
        group_by(Subplot_Descriptions, year) %>%
        summarize(avg_biomass = mean(Biomass, na.rm = TRUE))

# Plotting data
ggplot(legumes_avg, aes(x = year, y = avg_biomass, fill = Subplot_Descriptions)) +
        geom_jitter(shape=16, position=position_jitterdodge(), alpha = 0.6, aes(colour = Subplot_Descriptions)) +
        #scale_color_manual(values = c("gall" = "olivedrab", "no_gall" = "olivedrab3")) +
        geom_bar(position="dodge", stat = "identity") +
        labs(x = "Treatment", y = "Biomass (g)") +
        #scale_fill_manual(values = c("olivedrab", "olivedrab2")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "drought_insecticide" = "Drought \n Insecticide",
                                  "insecticide" = "Insecticide",
                                  "irrigated" = "Irrigated \n Control",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed \n Drought",
                                  "warmed_drought_insecticide" = "Warmed \n Drought \n Insecticide",
                                  "warmed_insecticide" = "Warmed \n Insecticide")) +
        theme(legend.position = "none")
```


```{r}
# select only red clover (Trpr) from ANPP dataframe = TRFPR
Trpr <- anpp_all[(anpp_all$Code == "TRFPR"),]

# Take subplot average of biomass
Trpr_avg <- Trpr %>%
        group_by(Subplot_Descriptions, Code) %>%
        summarise(avg_biomass = mean(Biomass, na.rm = TRUE),
        se = std.error(Biomass, na.rm = TRUE))

# select only red clover (Trpr) from ANPP dataframe = TRFPR
trpr21 <- anpp21[(anpp21$Code == "TRFPR"),]

# Take subplot average of biomass
trpr_avg21 <- trpr21 %>%
        group_by(Subplot_Descriptions, Code) %>%
        summarise(avg_biomass = mean(Biomass, na.rm = TRUE),
        se = std.error(Biomass, na.rm = TRUE))
```

```{r}
# select only red clover (Trpr) from ANPP dataframe = TRFPR
soca <- anpp21[(anpp21$Code == "SOOCA"),]

# Take subplot average of biomass
soca_avg21 <- soca %>%
        group_by(Subplot_Descriptions, Code) %>%
        summarise(avg_biomass = mean(Biomass, na.rm = TRUE),
        se = std.error(Biomass, na.rm = TRUE))
```


```{r}
# Plotting data
ggplot(Trpr_avg, aes(x = Subplot_Descriptions, y = avg_biomass, fill = Subplot_Descriptions)) +
        geom_jitter(shape=16, position=position_jitterdodge(), alpha = 0.6, aes(colour = Subplot_Descriptions)) +
        #scale_color_manual(values = c("gall" = "olivedrab", "no_gall" = "olivedrab3")) +
        geom_bar(stat = "identity") +
        labs(x = "Treatment", y = "Biomass (g)") +
        #scale_fill_manual(values = c("olivedrab", "olivedrab2")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "drought_insecticide" = "Drought \n Insecticide",
                                  "insecticide" = "Insecticide",
                                  "irrigated" = "Irrigated \n Control",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed \n Drought",
                                  "warmed_drought_insecticide" = "Warmed \n Drought \n Insecticide",
                                  "warmed_insecticide" = "Warmed \n Insecticide")) +
        theme(legend.position = "none")
```

Trpr minus the insecticide plots
```{r}
# removing treatments we aren't interested in
trpr_noninsecticide <- Trpr %>% 
  filter(!(Subplot_Descriptions=='drought_insecticide' |
             Subplot_Descriptions=='insecticide' |
             Subplot_Descriptions=='warmed_drought_insecticide' |
             Subplot_Descriptions=='warmed_insecticide'))

trpr_noninsecticide_avg <- Trpr_avg %>%
  filter(!(Subplot_Descriptions=='drought_insecticide' |
             Subplot_Descriptions=='insecticide' |
             Subplot_Descriptions=='warmed_drought_insecticide' |
             Subplot_Descriptions=='warmed_insecticide'))

# bar plot
png("trpr_noninsect_anpp_2021.png", units="in", width=6, height=6, res=300)
ggplot(trpr_noninsecticide_avg, aes(x = Subplot_Descriptions, y = avg_biomass, fill = Subplot_Descriptions)) +
        geom_jitter(shape=16, position=position_jitterdodge(), alpha = 0.6, aes(colour = Subplot_Descriptions)) +
        geom_bar(stat = "identity") +
        geom_errorbar(aes(ymin = avg_biomass - se, ymax = avg_biomass + se), width = 0.2,
                         position = "identity") +
        scale_fill_manual(values = c("ambient" = "darkblue", "drought" = "gray58", 
        "irrigated_control" = "deepskyblue", "warmed" = "red2", "warmed_drought" = "darkred")) +
        labs(x = "Treatment", y = "2021 T. pratense Biomass (g/0.20m^2)") +
        scale_x_discrete(limits = level_order,
                         labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "irrigated_control" = "Irrigated \n Control",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed \n Drought")) +
        theme(legend.position = "none")
dev.off
```

Soca plots
```{r}
# removing treatments we aren't interested in
soca_noninsecticide <- soca %>% 
  filter(!(Subplot_Descriptions=='drought_insecticide' |
             Subplot_Descriptions=='insecticide' |
             Subplot_Descriptions=='warmed_drought_insecticide' |
             Subplot_Descriptions=='warmed_insecticide'))

soca_noninsecticide_avg <- soca_avg21 %>%
  filter(!(Subplot_Descriptions=='drought_insecticide' |
             Subplot_Descriptions=='insecticide' |
             Subplot_Descriptions=='warmed_drought_insecticide' |
             Subplot_Descriptions=='warmed_insecticide'))

# bar plot
png("soca_noninsect_anpp_2021.png", units="in", width=6, height=6, res=300)
ggplot(soca_noninsecticide_avg, aes(x = Subplot_Descriptions, y = avg_biomass, fill = Subplot_Descriptions)) +
        geom_jitter(shape=16, position=position_jitterdodge(), alpha = 0.6, aes(colour = Subplot_Descriptions)) +
        geom_bar(stat = "identity") +
        geom_errorbar(aes(ymin = avg_biomass - se, ymax = avg_biomass + se), width = 0.2,
                         position = "identity") +
        scale_fill_manual(values = c("ambient" = "darkblue", "drought" = "gray58", 
        "irrigated_control" = "deepskyblue", "warmed" = "red2", "warmed_drought" = "darkred")) +
        labs(x = "Treatment", y = "Biomass (g/0.20m^2)") +
        scale_x_discrete(limits = level_order,
                         labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "irrigated_control" = "Irrigated \n Control",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed \n Drought")) +
        theme(legend.position = "none")
dev.off
```
```{r}
soca_trpr <- full_join(soca, trpr21)

soca_trpr_avg <- full_join(soca_avg21, trpr_avg21)

# removing treatments we aren't interested in
soca_trpr_noninsecticide <- soca_trpr %>% 
  filter(!(Subplot_Descriptions=='drought_insecticide' |
             Subplot_Descriptions=='insecticide' |
             Subplot_Descriptions=='warmed_drought_insecticide' |
             Subplot_Descriptions=='warmed_insecticide'))

soca_trpr_noninsecticide_avg <- soca_trpr_avg %>% 
  filter(!(Subplot_Descriptions=='drought_insecticide' |
             Subplot_Descriptions=='insecticide' |
             Subplot_Descriptions=='warmed_drought_insecticide' |
             Subplot_Descriptions=='warmed_insecticide'))

png("ANPP_L2_barplot_Trpr_Soca.png", units="in", width=6, height=6, res=300)
ggplot(soca_trpr_noninsecticide_avg, aes(x = Subplot_Descriptions, y = avg_biomass, fill = Code)) +
        geom_bar(position = "dodge", stat = "identity", col = "black") +
        geom_errorbar(aes(ymin = avg_biomass - se, ymax = avg_biomass + se), width = 0.2,
                      position = position_dodge(0.9)) +
        labs(x = "Treatment", y = bquote("2021 Average Biomass g/0.20m^2"), fill="Species") +
        scale_x_discrete(limits = level_order,
                         labels=c("ambient" = "Ambient",
                                  "warmed" = "Warmed",
                                  "irrigated_control" = "Irrigated \n Control",
                                  "drought" = "Drought",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        scale_fill_manual(values=c("#edae49", "#d1495b"),
                          labels=c("Solidago canadensis","Trifolium pratense")) +
        theme(legend.position = "top")
dev.off
```

Data Exploration
```{r}
# checking raw data
hist(trpr_noninsecticide$Biomass)
qqnorm(trpr_noninsecticide$Biomass)
shapiro.test(trpr_noninsecticide$Biomass) # pvalue is > 0.05 so we can reject the null hypothesis that the data is normal (aka normally distributed)

# Exploring distributions for these data:
descdist(trpr_noninsecticide$Biomass, discrete = FALSE)

# checking model 
# is it normally distributed? yes
m1 <- lm(Biomass ~ Subplot_Descriptions, data = trpr_noninsecticide)
hist(resid(m1))
qqnorm(resid(m1))
plot(m1)
# homogeneity of variance? yes, p >0.05 (no significant difference between the group variances)
leveneTest(residuals(m1) ~ trpr_noninsecticide$Subplot_Descriptions)

### Model exploration ###
m2 <- lmer(Biomass ~ Subplot_Descriptions + (1|Replicate/Footprint_Location), data=trpr_noninsecticide, REML=F)
summary(m2)
emmeans(m2, list(pairwise ~ Subplot_Descriptions), adjust = "tukey")
```


Data Exploration
```{r}
# checking raw data
hist(soca_noninsecticide$Biomass)
qqnorm(soca_noninsecticide$Biomass)
shapiro.test(soca_noninsecticide$Biomass) # pvalue is > 0.05 so we can reject the null hypothesis that the data is normal (aka normally distributed)

# Exploring distributions for these data:
descdist(soca_noninsecticide$Biomass, discrete = FALSE)

# checking model 
# is it normally distributed? yes
m1s <- lm(Biomass ~ Subplot_Descriptions, data = soca_noninsecticide)
hist(resid(m1s))
qqnorm(resid(m1s))
plot(m1s)
# homogeneity of variance? yes, p >0.05 (no significant difference between the group variances)
leveneTest(residuals(m1s) ~ soca_noninsecticide$Subplot_Descriptions)

### Model exploration ###
m2s <- lmer(Biomass ~ Subplot_Descriptions + (1|Replicate/Footprint_Location), data=soca_noninsecticide, REML=F)
summary(m2s)
emmeans(m2s, list(pairwise ~ Subplot_Descriptions), adjust = "tukey")
```


Data Exploration all biomass - 2021
```{r}
# checking raw data
hist(anpp21$Biomass)
qqnorm(anpp21$Biomass)
shapiro.test(anpp21$Biomass) # pvalue is > 0.05 so we can reject the null hypothesis that the data is normal (aka normally distributed)

# Exploring distributions for these data:
descdist(anpp21$Biomass, discrete = FALSE)

# checking model 
# is it normally distributed? yes
m1 <- lm(log(Biomass) ~ Subplot_Descriptions, data = anpp21)
hist(resid(m1))
qqnorm(resid(m1))
plot(m1)
# homogeneity of variance? yes, p >0.05 (no significant difference between the group variances)
leveneTest(residuals(m1) ~ anpp21$Subplot_Descriptions)

### Model exploration ###
m2 <- lmer(Biomass ~ Subplot_Descriptions + (1|Replicate/Footprint_Location), data=anpp21, REML=F)
summary(m2)
emmeans(m2, list(pairwise ~ Subplot_Descriptions), adjust = "tukey")
```

Data Exploration all biomass - 2021
```{r}
# remove species we aren't interested in (i.e., unknowns and unsorted plant material)
anpp21_1 <- anpp21 %>%
  filter(!(Code=='UNASTAE' |
           Code=='UNDIC' |
           Code=='UNFABAE' |
           Code=='UNSRT'))

# checking raw data
hist(anpp21_1$Biomass)
qqnorm(anpp21_1$Biomass)
shapiro.test(anpp21_1$Biomass) # pvalue is > 0.05 so we can reject the null hypothesis that the data is normal (aka normally distributed)

# Exploring distributions for these data:
descdist(anpp21_1$Biomass, discrete = FALSE)

# checking model 
# is it normally distributed? yes
m1 <- lm(log(Biomass) ~ Subplot_Descriptions, data = anpp21)
hist(resid(m1))
qqnorm(resid(m1))
plot(m1)
# homogeneity of variance? yes, p >0.05 (no significant difference between the group variances)
leveneTest(residuals(m1) ~ anpp21_1$Subplot_Descriptions)

### Model exploration ###
m2 <- lmer(Biomass ~ Subplot_Descriptions + (1|Replicate/Footprint_Location), data=anpp21_1, REML=F)
summary(m2)
emmeans(m2, list(pairwise ~ Subplot_Descriptions), adjust = "tukey")
```