---
title: "REX: T7 Red Clover ANPP Analyses"
author: "Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
---

COLLABORATORS: Kara Dobson, Phoebe Zarnetske, Mark Hammond
DATA INPUT:  csv from the shared Google drive  
DATA OUTPUT: Code and Rmd are in the scripts folder in Github  
PROJECT: REX  

DESCRIPTION:
- Looking at legume species (primarily red clover aka Trifolium pratense - Trpr)
- Main question: have the treatments (warming & drought) altered the amount of biomass and/or within certain species such as legumes?
- 2019 is baseline data taken by Mark Hammond before any treatments were imposed (1m2)
- 2021 is data from the first year of REX (0.25m2)

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

# Load packages
library(bbmle)
library(lmerTest)
library(fitdistrplus)
library(sjPlot)
library(car)
library(emmeans)
library(tidyverse)
library(ggpubr)
library(jtools) #summ 
library(rstatix)

# Set working directory from .Renviron
dir <- Sys.getenv("DATA_DIR")
list.files(dir)

# Read in data
ANPP_21 <- read.csv(file.path(dir, "T7_ANPP/L1/T7_Zarnetske_ANPP_L1.csv")) # 2021 ANPP data
ANPP_19 <- read.csv(file.path(dir, "T7_ANPP/L0/REX_T7_ANPP_all_plots_2019.csv")) # not cleaned data - 2019 data
meta <- read.csv(file.path(dir, "REX_warmx_metadata.csv"))
```

Looking at all ANPP from 2019 - baseline data
```{r}
# cleaning data
anpp_19 <- ANPP_19 %>% select(Date, PZ_footprints_Field_Code, Group_Code, Dried_plant_biomass_g_per_metersquare)
names(anpp_19)[2] <- "Unique_ID" #changing column name to match what's in the meta data file
anpp_19_merge <- merge(anpp_19, meta)

# select only legume data from ANPP 2019
legume_19 <- anpp_19_merge[(anpp_19_merge$Group_Code == "Legume"),]

# Take treatment average total biomass
legume_19_avg <- legume_19 %>%
        group_by(Subplot_Descriptions) %>%
        summarize(avg_biomass = mean(Dried_plant_biomass_g_per_metersquare, na.rm = TRUE))

# Plotting data
ggplot(legume_19_avg, aes(x = Subplot_Descriptions, y = avg_biomass, fill = Subplot_Descriptions)) +
        geom_jitter(shape=16, position=position_jitterdodge(), alpha = 0.6, aes(colour = Subplot_Descriptions)) +
        #scale_color_manual(values = c("gall" = "olivedrab", "no_gall" = "olivedrab3")) +
        geom_bar(stat = "identity") +
        labs(x = "Treatment", y = "Biomass (g)") +
        #scale_fill_manual(values = c("olivedrab", "olivedrab2")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "drought_insecticide" = "Drought \n Insecticide",
                                  "insecticide" = "Insecticide",
                                  "irrigated" = "Irrigated \n Control",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed \n Drought",
                                  "warmed_drought_insecticide" = "Warmed \n Drought \n Insecticide",
                                  "warmed_insecticide" = "Warmed \n Insecticide")) +
        theme(legend.position = "none")
```

Looking at all ANPP from 2020
```{r}
# Take subplot total biomass
anpp_sum <- ANPP_21 %>%
        group_by(Field_Loc_Code, Subplot_Descriptions) %>%
        summarize(sum_biomass = sum(Dried_Plant_Biomass_g, na.rm = TRUE))

# Take treatment average total biomass
anpp_avg <- anpp_sum %>%
        group_by(Subplot_Descriptions) %>%
        summarize(avg_biomass = mean(sum_biomass, na.rm = TRUE))

# Plotting data from average total biomass of each treatment
ggplot(anpp_avg, aes(x = Subplot_Descriptions, y = avg_biomass, fill = Subplot_Descriptions)) +
        geom_jitter(shape=16, position=position_jitterdodge(), alpha = 0.6, aes(colour = Subplot_Descriptions)) +
        #scale_color_manual(values = c("gall" = "olivedrab", "no_gall" = "olivedrab3")) +
        geom_bar(stat = "identity") +
        labs(x = "Treatment", y = "Biomass (g)") +
        #scale_fill_manual(values = c("olivedrab", "olivedrab2")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "drought_insecticide" = "Drought \n Insecticide",
                                  "insecticide" = "Insecticide",
                                  "irrigated" = "Irrigated \n Control",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed \n Drought",
                                  "warmed_drought_insecticide" = "Warmed \n Drought \n Insecticide",
                                  "warmed_insecticide" = "Warmed \n Insecticide")) +
        theme(legend.position = "none")
```
Looking at all legume species (Fabaceae family)
```{r}
# select only Fabaceae from ANPP 2021 dataframe
Fabaceae <- ANPP_21[(ANPP_21$family == "Fabaceae"),]
Fabaceae <- na.omit(Fabaceae) # remove NAs
```

Looking only at red clover (Trpr or TRFPR)
```{r}
# select only red clover (Trpr) from ANPP dataframe = TRFPR
Trpr <- ANPP_21[(ANPP_21$Species_Code == "TRFPR"),] 

# multiply biomass values by 4 - biomass was only clipped at 0.25m2 of the 1m2 plot
Trpr$biomass_scaled <- Trpr$Dried_Plant_Biomass_g*4

# Take subplot average of biomass
Trpr_avg_scaled <- Trpr %>%
        group_by(Subplot_Descriptions) %>%
        summarize(avg_biomass_scaled = mean(biomass_scaled, na.rm = TRUE))

# Take subplot average of biomass
Trpr_avg <- Trpr %>%
        group_by(Subplot_Descriptions) %>%
        summarize(avg_biomass = mean(Dried_Plant_Biomass_g, na.rm = TRUE))
```

```{r}
# Plotting data
ggplot(Trpr_avg, aes(x = Subplot_Descriptions, y = avg_biomass, fill = Subplot_Descriptions)) +
        geom_jitter(shape=16, position=position_jitterdodge(), alpha = 0.6, aes(colour = Subplot_Descriptions)) +
        #scale_color_manual(values = c("gall" = "olivedrab", "no_gall" = "olivedrab3")) +
        geom_bar(stat = "identity") +
        labs(x = "Treatment", y = "Biomass (g)") +
        #scale_fill_manual(values = c("olivedrab", "olivedrab2")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "drought_insecticide" = "Drought \n Insecticide",
                                  "insecticide" = "Insecticide",
                                  "irrigated" = "Irrigated \n Control",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed \n Drought",
                                  "warmed_drought_insecticide" = "Warmed \n Drought \n Insecticide",
                                  "warmed_insecticide" = "Warmed \n Insecticide")) +
        theme(legend.position = "none")
```

```{r}

```




