---
title: "Biolog Analysis Script"
author: "Adrian Noecker, Moriah Young"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
COLLABORATORS: Phoebe Zarnetske, Mark Hammond, Taylor Ulbrich (using code from Taylor)
DATA INPUT: L1 csv from the shared Google drive  
DATA OUTPUT: L2 csv and Code and Rmd are in the scripts folder in Github  
PROJECT: REX
---

TO DO:
- Make one data frame that includes all the diversity indices for all time points and upload to L2 folder
- Data analysis for Richness and AWCD

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

#load packages
library(car)
library(tidyr)
library(dplyr)
library(vegan)
library(tibble)
library(plyr)
library(readr)
library(tidyverse)
library(ggplot2)
library(lsmeans)
library(lme4)
library(tibble)
library(ggpubr)
library(lmerTest)
library(fitdistrplus)
library(sjPlot)
library(tidyverse)
library(car)
library(emmeans)
library(bbmle)
library(multcomp)

# Set working directory from .Renviron
dir <- Sys.getenv("DATA_DIR")
list.files(dir)
setwd(dir)

## Read in data
biolognowater <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/Biolog_data_NOWater_L1.csv"))
standardbiolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/Biolog_data_L1.csv"))

# Moriah's below (for some reason above wont work for me - not same file path)
biolognowater <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/Biolog_data_NOWater_L1.csv"))
standardbiolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/Biolog_data_L1.csv"))
```

#reformatting and checking absorbances
# Moriah - this part is a little confusing. For example, the differentiation between the standard and biolog dataframes
```{r}
#removing water from standard biolog
standardnowater <- standardbiolog %>% filter(!Well_C_Source %in% c("Water"))

#changing row name and reformatting, combine hour and ID to one column
biologdf <- biolognowater %>%
  unite(SampleID_Hour, Biolog_column_ID, Hour, sep ="_", remove = FALSE)

standarddf <- standardnowater %>%
  unite(SampleID_Hour, Biolog_column_ID, Hour, sep ="_", remove = FALSE)

#renaming new column
biologdf$rowname <- biologdf$SampleID_Hour

# Note from Moriah: is there a reason why the first column is converted to the row names??
#changing column to row
biologmatrix <- column_to_rownames(biologdf, "rowname")

#Check control absorbances (all should == 0 if no contamination)
biologcontrol <- biologmatrix %>% filter(Sample_ID=="BIOLOG CONTROL")
####Tween 40, Tween 80, D Xylose, X4 Hydroxy Benzoic Acid, and X2 Hydroxy Benzoic Acid larger than 0.01 in a few controls, most hour 0 in plates 5, 7, 9, and 10 so maybe just the initial color of well and not contamination??

standardcontrol <- standarddf %>% filter(Sample_ID=="BIOLOG CONTROL")

#exclude controls for analysis
biologsamples <- biologmatrix %>% filter(!Sample_ID %in% c("BIOLOG CONTROL"))

standardsamples <- standarddf %>% filter(!Sample_ID %in% c("BIOLOG CONTROL"))

#import meta data to get treatment type
meta <- read.csv(file.path(dir, "/REX_warmx_metadata.csv"))
                 
#get rid of unneeded columns
meta1 <- meta %>% dplyr::select(Subplot_Descriptions, Unique_ID, Rep, Footprint_Location)

# Note from Moriah: when we merge these two together, we lose the matrix formatting - may need to run "column_to_rownames" function again on new dataframe to make SampleID_Hour row names
#combine meta and sample df
biologsamples1 <- merge(x=biologsamples, y=meta1, by.x=c('Plot_ID'), 
      by.y=c('Unique_ID'))

standardsamples1 <- merge(x=standardsamples, y=meta1, by.x=c('Plot_ID'), 
      by.y=c('Unique_ID'))

#change column order
biologsamples2 <- biologsamples1 %>% relocate(Subplot_Descriptions, .after = Biolog_column_ID)

```

#grouping by hour
```{r}
# split data by Hour so we can calculate AWCD by each Hour 
biologHr0 <- filter(biologsamples2, Hour == 0)
biologHr24 <- filter(biologsamples2, Hour == 24)
biologHr48 <- filter(biologsamples2, Hour == 48)
biologHr72 <- filter(biologsamples2, Hour == 72)
biologHr96 <- filter(biologsamples2, Hour == 96)
biologHr120 <- filter(biologsamples2, Hour == 120)
biologHr144 <- filter(biologsamples2, Hour == 144)
biologHr168 <- filter(biologsamples2, Hour == 168)
```

#Diversity Calculations seperated by hour
```{r}
###Hour 168
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness168hr <- rowSums(biologHr168[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon168hr <- diversity(biologHr168[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD168hr <- apply(biologHr168[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr168diversity <- cbind(biologHr168[1:6], Richness168hr, Shannon168hr, AWCD168hr)

###Hour 144
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness144hr <- rowSums(biologHr144[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon144hr <- diversity(biologHr144[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD144hr <- apply(biologHr144[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr144diversity <- cbind(biologHr144[1:6], Richness144hr, Shannon144hr, AWCD144hr)

###Hour 120
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness120hr <- rowSums(biologHr120[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon120hr <- diversity(biologHr120[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD120hr <- apply(biologHr120[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr120diversity <- cbind(biologHr120[1:6], Richness120hr, Shannon120hr, AWCD120hr)

###Hour 96
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness96hr <- rowSums(biologHr96[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon96hr <- diversity(biologHr96[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD96hr <- apply(biologHr96[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr96diversity <- cbind(biologHr96[1:6], Richness96hr, Shannon96hr, AWCD96hr)

###Hour 72
# First, richness of subsrates used; add all indices as a column to the metadata file 
Richness72hr <- rowSums(biologHr72[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon72hr <- diversity(biologHr72[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD72hr <- apply(biologHr72[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr72diversity <- cbind(biologHr72[1:6], Richness72hr, Shannon72hr, AWCD72hr)

###Hour 48
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness48hr <- rowSums(biologHr48[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon48hr <- diversity(biologHr48[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD48hr <- apply(biologHr48[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr48diversity <- cbind(biologHr48[1:6], Richness48hr, Shannon48hr, AWCD48hr)

###Hour 24
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness24hr <- rowSums(biologHr24[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon24hr <- diversity(biologHr24[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD24hr <- apply(biologHr24[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr24diversity <- cbind(biologHr24[1:6], Richness24hr, Shannon24hr, AWCD24hr)

```

# PLOTS

# Diversity plots
# Hour 168 - day 7
```{r}
# RICHNESS 
ggplot(biologHr168diversity, aes(x = Subplot_Descriptions, y = Richness168hr, fill = Subplot_Descriptions)) +
        geom_boxplot() + 
        labs(x = "Treatment", 
             y = "Catabolic Richness", 
             title = "Microbial Catabolic Richness, Day 7") + 
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

# SHANNON DIVERSITY 
ggplot(biologHr168diversity, aes(x = Subplot_Descriptions, y = Shannon168hr, fill = Subplot_Descriptions)) +
        geom_boxplot() + 
        labs(x = "Treatment", 
             y = "Catabolic Shannon Diversity", 
             title = "Microbial Catabolic Shannon Diversity, Day 7") + 
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

# AWCD
ggplot(biologHr168diversity, aes(x = Subplot_Descriptions, y = AWCD168hr, fill = Subplot_Descriptions)) +
        geom_boxplot() + 
        labs(x = "Treatment", 
             y = "Catabolic AWCD", 
             title = "Microbial Catabolic AWCD, Day 7") + 
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

```

Plots - looking at individual carbon sources and plotting absorbance levels across time (from time 0 - hr 168)
```{r}
#group by hour and sample for average
biolog2<- biologsamples[,!(names(biologsamples) %in% c("Plot_ID", "SampleID_Hour", "Sample_ID", "Biolog_column_ID"))]

##X2.Hydroxybenzoic Acid
#take average of each treatment per hour 
X2HydroxyBen_mean <- aggregate(biolog2$X2.Hydroxy_Benzoic_Acid, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("X2HydroxyBen_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(X2HydroxyBen_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="X2.Hydroxybenzoic Acid", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

##X4.Hydroxybenzoic Acid
#take average of each treatment per hour 
X4HydroxyBen_mean <- aggregate(biolog2$X4.Hydroxy_Benzoic_Acid, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("X4HydroxyBen_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(X4HydroxyBen_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="X4.Hydroxybenzoic Acid", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

##D.Cellobiose 
#take average of each treatment per hour 
DCell_mean <- aggregate(biolog2$D.Cellobiose, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("DCell_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(DCell_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="D.Cellobiose", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

##D.Galactonic Acid Y Lactone
#take average of each treatment per hour 
DGalactonic_mean <- aggregate(biolog2$D.Galactonic_Acid_γ.Lactone, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("DGalactonic_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(DGalactonic_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="D.Galactonic Acid Y Lactone", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

##D.Galacturonic Acid
#take average of each treatment per hour 
DGalacturonic_mean <- aggregate(biolog2$D.Galacturonic_Acid, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("DGalacturonic_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(DGalacturonic_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="D.Galacturonic Acid", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

```

# DATA ANALYSIS

Notes from Moriah:
- Model structure is a little more complex in REX - need to have random effects as such: (1|Rep/Footprint/Subplot). This is a nested plot design. For lack of better explanation, it goes from largest "scale" (rep) to smallest "scale" (subplot) This means that you'll want to retain the rep and footprint numbers in the dataframe you ultimately create below and with the diversity metrics.

#Looking at distribution of data
```{r}
###Starting with final reading (hour=168) 

# test for normal residuals and equal variances (ANOVA assumptions)
# normal residuals?
ggqqplot(biologHr168diversity$Richness168hr)
hist(biologHr168diversity$Richness168hr) ##slightly skewed left

# Shapiro test - null hypothesis: residuals are normally distributed 
require(lme4)
require(lmerTest)
model = lm(Richness168hr ~ Subplot_Descriptions, data = biologHr168diversity)
res = model$residuals
shapiro.test(res) 
#p = 0.002237 so not normally distributed?

# homogeneity of variances
# Many statistical tests (like a one-way ANOVA) assume that variances are equal across samples. Bartlett’s test can be 
# used to verify that assumption.
# H0: The variance among each group is equal.
# HA: At least one group has a variance that is not equal to the rest.

bartlett.test(Richness168hr ~ Subplot_Descriptions, data = biologHr168diversity) 
# 0.6396 = we fail to reject the null hypothesis, so all groups have the same variance. This is what we want.

#ANOVA
Anova(model, type = "III")

#Response: Richness for hour 168
#                     Sum Sq Df F value    Pr(>F)    
#(Intercept)          3750.0  1 79.9290 2.009e-08 ***
#Subplot_Descriptions   53.0  3  0.3766    0.7709    
#Residuals             938.3 20   

# confidence intervals for every pairwise comparison
tukey <- TukeyHSD(aov(Richness168hr ~ Subplot_Descriptions, data = biologHr168diversity))
tukey
plot(tukey)

#Subplot_Descriptions
#                            diff        lwr       upr     p adj
#drought-ambient        -0.500000 -11.568683 10.568683 0.9992524
#warmed-ambient         -3.833333 -14.902016  7.235349 0.7680882
#warmed_drought-ambient -1.000000 -12.068683 10.068683 0.9941408
#warmed-drought         -3.333333 -14.402016  7.735349 0.8334310
#warmed_drought-drought -0.500000 -11.568683 10.568683 0.9992524
#warmed_drought-warmed   2.833333  -8.235349 13.902016 0.8893834

```

Example script from other REX data on data analysis with steps for checking for assumption
https://github.com/kbs-lter/rex-analysis/blob/main/R/T7_warmx_plant_traits/Soca_plant_height_analyses_L2.R

# Checking Distribution of all reads across all hours
```{r}
#calculate shannon diversity for all reads
ShannonAll <- diversity(biologsamples2[,c(7:37)], index = "shannon")

#histogram of shannon index for each sample each hour all reads-skewed right
hist(ShannonAll)

#shapiro test
# normal: p > 0.05
# p=9.89e-5 - not normal
shapiro.test(ShannonAll)

#bind with rest of data
biolog.shannon <- cbind(biologsamples2, ShannonAll)

#look at final reads for lmer
biologshannon168 <- filter(biolog.shannon, Hour == 168)

```

# LMER 
```{r}
# without checking assumptions - need to do this. Just laying out the mixed effects model
m2 <- lmer(ShannonAll ~ Subplot_Descriptions + (1|Rep/Footprint_Location), data = biologshannon168, REML=F)
summary(m2) #looking at just hour 168 is singular
plot(m2, main = ShannonAll)

#levene test for homogeneity of variance
leveneTest(residuals(m2) ~ biologshannon168$Subplot_Descriptions) 

# Check for normal residuals
qqPlot(resid(m2), main = "ShannonAll")
hist(residuals(m2), main = "ShannonAll")
shapiro.test(resid(m2)) #p=0.02 - not normal but close

#anova
anova(m2)

```

## WORK IN PROGRESS!!
# Need to go back to when the matrices were made for each our and make sure the "subplot_descriptions" are included
## PCA with arrows
```{r}
# select only for hour 168
biolog.samples.Hr168 <- filter(biologsamples2, Hour == "168")

# make sample names the rownames so we keep them during ordination 
biolog.matrix.Hr168 <- column_to_rownames(biolog.samples.Hr168, "Plot_ID")


############ Multivariate analysis ##################
map<-biolog.matrix.Hr168[,1:5] #Converting  data to two matrices: Mapping and substrate (Comm)
comm<-biolog.matrix.Hr168[,6:35]
comm<-comm/rowSums(comm) #Relative abundance

# distance of rel. abund. matrix
dist<-vegdist(comm, method='bray', na.rm=TRUE)

#Principal component analysis
apcoa<-cmdscale(dist, k=2, eig=FALSE, add=FALSE, x.ret=FALSE)
plot<-ordiplot(apcoa, type="n")
vecpcoa<-envfit(apcoa,comm)
plot(vecpcoa)# add vectors
vecpcoa$vectors

# Set the colors by treatment 

points(plot, "sites", pch = 20, col = "green4", map$treatment=="SL_0") 
points(plot, "sites", pch = 20, col = "chartreuse",map$treatment=="SL_SL") 
points(plot, "sites", pch = 20, col = "blue1",map$treatment=="SL_BBS") 
points(plot, "sites", pch = 20, col = "darkmagenta",map$treatment=="SL_JG")
points(plot, "sites", pch = 20, col = "gold",map$treatment=="SL_S") 


plot + theme(legend.text = element_text(size = 16))

```
