---
title: "Biolog 2022 Plots"
author: "Adrian Noecker, Moriah Young"
date: "September 2023, updated October 2024"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
COLLABORATORS: Phoebe Zarnetske, Mark Hammond, Taylor Ulbrich (using code from Taylor)
DATA INPUT: L1 csv from the shared Google drive  
DATA OUTPUT: L2 csv and Code and Rmd are in the scripts folder in Github  
PROJECT: REX
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

#load packages
library(car)
library(tidyr)
library(plotrix)
library(dplyr)
library(tibble)
library(plyr)
library(readr)
library(tidyverse)
library(ggplot2)
library(tibble)
library(ggpubr)
library(fitdistrplus)
library(sjPlot)
library(tidyverse)
library(multcomp)
library(vegan)

# Set working directory from .Renviron
dir <- Sys.getenv("DATA_DIR")
list.files(dir)

## Read in data
biolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/biolog_clean_wide_format_L1.csv"))
diversity <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L2/biolog_hill_diversity_L2.csv"))
biolog2 <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/biolog_clean_L1.csv"))

# Moriah's below (for some reason above wont work for me - not same file path)
biolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/biolog_clean_wide_format_L1.csv"))
biolog2 <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/biolog_clean_L1.csv"))
diversity <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L2/biolog_hill_diversity_L2.csv"))
```

Maybe move this to cleaning script
Grouping carbon sources
```{r}
carbon_groups <- list(
        "amino_acids" = c("β-HydroxyGlycyl-L-Glutamic_Acid", "L-Threonine", "L-Serine", "L-Asparagine", "L-Phenylalanine", "L-Arginine"),
        "amines" = c("Putrescine", "Phenylethylamine"),
        "carbohydrates" = c("β-MethylD-Glucoside", "D-Xylose", "i-Erythritol", "D-Mannitol", "N-Acetyl-D-Glucosamine", "D-Cellobiose", "Glucose1-Phosphate", "α-D-Lactose", "D,L-α-Glycerol_Phosphate"),
        "carboxylic_acids" = c("D-Malic_Acid", "α-Keto_Butyric_Acid", "Itaconic_Acid", "Pyruvic_Acid_Methyl_Ester", "γ-Amino_Butyric_Acid", "D-Galactonic_Acid_γ-Lactone", "D-Glucosaminic_Acid", "D-Galacturonic_Acid"),
        "phenolic_acids" = c("4-Hydroxy_Benzoic_Acid", "2-Hydroxy_Benzoic_Acid"),
        "polymers" = c("Glycogen", "Tween_40", "Tween_80", "α-Cyclodextrin")
        )

# Create the new column with broader group names based on the mapping
biolog2$carbon_group <- unlist(sapply(biolog2$Well_C_Source, function(cs) {
  group <- names(carbon_groups)[sapply(carbon_groups, function(x) cs %in% x)]
  if (length(group) == 0) group <- "Unknown"
  group
}))

# move Absorbence column to be the last column
biolog2 <- biolog2 %>%
  relocate(Absorbence, .after = last_col())

# Calculate mean and se for absorbence by subplot descriptions, carbon group, and hour
# Define a custom function to calculate mean and standard error
calculate_mean_se <- function(x) {
  mean_val <- mean(x, na.rm = TRUE)
  se_val <- sd(x, na.rm = TRUE) / sqrt(length(x))
  return(c(mean_val, se_val))
}

# Apply the custom function with aggregate for mean
mean_result <- aggregate(Absorbence ~ Subplot_Descriptions + carbon_group + Hour, data = biolog2, FUN = function(x) mean(x, na.rm = TRUE))

# Apply the custom function with aggregate for standard error
se_result <- aggregate(Absorbence ~ Subplot_Descriptions + carbon_group + Hour, data = biolog2, FUN = function(x) sd(x, na.rm = TRUE) / sqrt(length(x)))

# Merge the results based on the grouping variables
result <- merge(mean_result, se_result, by = c("Subplot_Descriptions", "carbon_group", "Hour"))

# Rename the columns
colnames(result)[4:5] <- c("avg_absorbence", "se")

# Apply the custom function with aggregate
#avg_biolog <- aggregate(Absorbence ~ Subplot_Descriptions + carbon_group + Hour, data = biolog2, FUN = mean)
```

# carbon group plots
```{r}
# setting up for plots
carbon_groups1 <- c("amines" = "Amines",
                    "amino_acids" = "Amino Acids",
                    "carbohydrates" = "Carbohydrates",
                    "carboxylic_acids" = "Carboxylic Acids",
                    "phenolic_acids" = "Phenolic Acids",
                    "polymers" = "Polymers")

# Hour 168 - day 7
biologHr168 <- result %>% filter(result$Hour == "168")

# plot that shows average AWCD for each treatment faceted out by carbon group at hour 168
ggplot(biologHr168, aes(x = Subplot_Descriptions, y = avg_absorbence, fill = Subplot_Descriptions)) +
        geom_pointrange(aes(ymin=avg_absorbence-se, ymax=avg_absorbence+se), 
                        pch=21,
                        size=2,
                        position=position_dodge(0.3)) + 
        facet_wrap(~carbon_group, labeller = as_labeller(carbon_groups1)) +
        labs(x = "Treatment", 
             y = "Average AWCD", 
             title = "Average AWCD, Day 7") + 
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

# Hour 48 - day 2
biologHr48 <- result %>% filter(result$Hour == "48")

# plot that shows average AWCD for each treatment faceted out by carbon group at hour 168
ggplot(biologHr48, aes(x = Subplot_Descriptions, y = avg_absorbence, fill = Subplot_Descriptions)) +
        geom_pointrange(aes(ymin=avg_absorbence-se, ymax=avg_absorbence+se), 
                        pch=21,
                        size=1,
                        position=position_dodge(0.3)) + 
        facet_wrap(~carbon_group, labeller = as_labeller(carbon_groups1)) +
        labs(x = "Treatment", 
             y = "Average AWCD", 
             title = "Average AWCD, Day 2") + 
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

#Filtering for Each Subplot Description
ambient=as.data.frame(filter(result, Subplot_Descriptions=="ambient"))

drought=as.data.frame(filter(result, Subplot_Descriptions=="drought"))

warmed=as.data.frame(filter(result, Subplot_Descriptions=="warmed"))

warmeddrought=as.data.frame(filter(result, Subplot_Descriptions=="warmed_drought"))

#Plotting avg absorbance of each treatment across time
#png("AmbientAWCDbyCgroups", units="in", width=7, height=6, res=300)
ggplot(ambient, aes(x = Hour, y = avg_absorbence, color = carbon_group)) +
        labs(x = "Hour", 
             y = "AWCD", 
             title = "Ambient AWCD") +
  geom_line() + geom_point() +
 scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E", "forestgreen", "darkblue")) +
        theme(legend.position = "left",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))
#dev.off()

#png("WarmedAWCDbyCgroups", units="in", width=7, height=6, res=300)
ggplot(warmed, aes(x = Hour, y = avg_absorbence, color = carbon_group)) +
        labs(x = "Hour", 
             y = "AWCD", 
             title = "Warmed AWCD") +
  geom_line() + geom_point() +
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E", "forestgreen", "darkblue")) +
        theme(legend.position = "left",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))
#dev.off()

#png("DroughtAWCDbyCgroups", units="in", width=7, height=6, res=300)
ggplot(drought, aes(x = Hour, y = avg_absorbence, color = carbon_group)) +
        labs(x = "Hour", 
             y = "AWCD", 
             title = "Drought AWCD") +
  geom_line() + geom_point() +
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E", "forestgreen", "darkblue")) +
        theme(legend.position = "left",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))
#dev.off()

#png("WarmedDroughtAWCDbyCgroups", units="in", width=7, height=6, res=300)
ggplot(warmeddrought, aes(x = Hour, y = avg_absorbence, color = carbon_group)) +
        labs(x = "Hour", 
             y = "AWCD", 
             title = "Warmed + Drought AWCD") +
  geom_line() + geom_point() +
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E", "forestgreen", "darkblue")) +
        theme(legend.position = "left",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))
#dev.off()

#Sorting for each carbon group
amines=as.data.frame(filter(result, result$carbon_group=="amines"))
aminoacids=as.data.frame(filter(result, result$carbon_group=="amino_acids"))
carbs=as.data.frame(filter(result, result$carbon_group=="carbohydrates"))
carboxylicacids=as.data.frame(filter(result, result$carbon_group=="carboxylic_acids"))
phenolicacids=as.data.frame(filter(result, result$carbon_group=="phenolic_acids"))
polymers=as.data.frame(filter(result, result$carbon_group=="polymers"))

#carbon group and treatment graphs
#amines
ggplot(amines, aes(x = Hour, y = avg_absorbence, color = Subplot_Descriptions)) +  geom_line() + geom_point() +
        geom_pointrange(aes(ymin=avg_absorbence-se, ymax=avg_absorbence+se)) +
        labs(x = "Treatment", 
             y = "Average AWCD", 
             title = "Average Amine AWCD") + 
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "right",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

#amino acids
ggplot(aminoacids, aes(x = Hour, y = avg_absorbence, color = Subplot_Descriptions)) +  geom_line() + geom_point() +
        geom_pointrange(aes(ymin=avg_absorbence-se, ymax=avg_absorbence+se)) +
        labs(x = "Treatment", 
             y = "Average AWCD", 
             title = "Average Amino Acid AWCD") + 
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "right",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

#Carbohydrates
ggplot(carbs, aes(x = Hour, y = avg_absorbence, color = Subplot_Descriptions)) +  geom_line() + geom_point() +
        geom_pointrange(aes(ymin=avg_absorbence-se, ymax=avg_absorbence+se)) +
        labs(x = "Treatment", 
             y = "Average AWCD", 
             title = "Average Carbohydrate AWCD") + 
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "right",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

# same code as above but adds position_dodge so that error bars, etc. don't overlap as much
ggplot(carbs, aes(x = Hour, y = avg_absorbence, color = Subplot_Descriptions)) +  
    geom_line(position = position_dodge(width = 0.5)) + 
    geom_point(position = position_dodge(width = 0.5)) +
    geom_pointrange(aes(ymin = avg_absorbence - se, ymax = avg_absorbence + se), 
                    position = position_dodge(width = 1)) +
    labs(x = "Treatment", 
         y = "Average AWCD", 
         title = "Average Carbohydrate AWCD") + 
    scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
    scale_x_discrete(labels = c("ambient" = "Ambient",
                                "drought" = "Drought",
                                "warmed" = "Warmed",
                                "warmed_drought" = "Warmed +\n Drought")) +
    theme(legend.position = "right",
          plot.title = element_text(hjust = 0.5, size = 20),
          axis.title.y = element_text(size = 20),
          axis.text.y = element_text(size = 15),
          axis.title.x = element_text(size = 20),
          axis.text.x = element_text(size = 15))

#carboxylic acids
ggplot(carboxylicacids, aes(x = Hour, y = avg_absorbence, color = Subplot_Descriptions)) +  geom_line() + geom_point() +
        geom_pointrange(aes(ymin=avg_absorbence-se, ymax=avg_absorbence+se)) +
        labs(x = "Treatment", 
             y = "Average AWCD", 
             title = "Average Carboxylic Acid AWCD") + 
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "right",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

#Phenolic Acids
ggplot(phenolicacids, aes(x = Hour, y = avg_absorbence, color = Subplot_Descriptions)) +  
        geom_line() + 
        geom_point() +
        geom_pointrange(aes(ymin=avg_absorbence-se, ymax=avg_absorbence+se)) +
        labs(x = "Treatment", 
             y = "Average AWCD", 
             title = "Average Phenolic Acids AWCD") + 
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "right",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

#Polymers
ggplot(polymers, aes(x = Hour, y = avg_absorbence, color = Subplot_Descriptions)) +  geom_line() + geom_point() +
        geom_pointrange(aes(ymin=avg_absorbence-se, ymax=avg_absorbence+se)) +
        labs(x = "Treatment", 
             y = "Average AWCD", 
             title = "Average Polymer AWCD") + 
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "right",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))

```


Diversity plots
```{r}
# Hour 168 - day 7
biologHr168diversity <- diversity %>% filter(diversity$Hour == "168")

# RICHNESS 
png("richness168hr.png", units="in", width=7, height=6, res=300)
ggplot(biologHr168diversity, aes(x = Subplot_Descriptions, y = H0, fill = Subplot_Descriptions)) +
        geom_boxplot() + 
        labs(x = "Treatment", 
             y = "Catabolic Richness", 
             title = "Microbial Catabolic Richness, Day 7") + 
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))
dev.off()

# SHANNON DIVERSITY 
png("shannon168hr.png", units="in", width=7, height=6, res=300)
ggplot(biologHr168diversity, aes(x = Subplot_Descriptions, y = H1, fill = Subplot_Descriptions)) +
        geom_boxplot() + 
        labs(x = "Treatment", 
             y = "Catabolic Shannon Diversity", 
             title = "Microbial Catabolic Shannon Diversity, Day 7") + 
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))
dev.off()

# SIMPSON DIVERSITY 
png("shannon168hr.png", units="in", width=7, height=6, res=300)
ggplot(biologHr168diversity, aes(x = Subplot_Descriptions, y = H2, fill = Subplot_Descriptions)) +
        geom_boxplot() + 
        labs(x = "Treatment", 
             y = "Catabolic Simpson Diversity", 
             title = "Microbial Catabolic Shannon Diversity, Day 7") + 
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))
dev.off()

# AWCD
png("AWCD168hr.png", units="in", width=7, height=6, res=300)
ggplot(biologHr168diversity, aes(x = Subplot_Descriptions, y = avg_AWCD, fill = Subplot_Descriptions)) +
        geom_boxplot() + 
        labs(x = "Treatment", 
             y = "Catabolic AWCD", 
             title = "Microbial Catabolic AWCD, Day 7") + 
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E")) +
        scale_x_discrete(labels=c("ambient" = "Ambient",
                                  "drought" = "Drought",
                                  "warmed" = "Warmed",
                                  "warmed_drought" = "Warmed +\n Drought")) +
        theme(legend.position = "none",
              plot.title = element_text(hjust = 0.5, size = 20),
              axis.title.y = element_text(size = 20),
              axis.text.y = element_text(size = 15),
              axis.title.x = element_text(size = 20),
              axis.text.x = element_text(size = 15))
dev.off()
```

Plots - looking at individual carbon sources and plotting absorbency levels across time (from time 0 - hr 168)
```{r}
#group by hour and sample for average
biolog2 <- biolog[,!(names(biolog) %in% c("Plot_ID", "SampleID_Hour", "Sample_ID", "Biolog_column_ID"))]

##X2.Hydroxybenzoic Acid
#take average of each treatment per hour 
X2HydroxyBen_mean <- aggregate(biolog2$X2.Hydroxy_Benzoic_Acid, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("X2HydroxyBen_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(X2HydroxyBen_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="X2.Hydroxybenzoic Acid", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

##X4.Hydroxybenzoic Acid
#take average of each treatment per hour 
X4HydroxyBen_mean <- aggregate(biolog2$X4.Hydroxy_Benzoic_Acid, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("X4HydroxyBen_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(X4HydroxyBen_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="X4.Hydroxybenzoic Acid", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

##D.Cellobiose 
#take average of each treatment per hour 
DCell_mean <- aggregate(biolog2$D.Cellobiose, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("DCell_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(DCell_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="D.Cellobiose", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

##D.Galactonic Acid Y Lactone
#take average of each treatment per hour 
DGalactonic_mean <- aggregate(biolog2$D.Galactonic_Acid_γ.Lactone, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("DGalactonic_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(DGalactonic_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="D.Galactonic Acid Y Lactone", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

##D.Galacturonic Acid
#take average of each treatment per hour 
DGalacturonic_mean <- aggregate(biolog2$D.Galacturonic_Acid, by=list(biolog2$Hour,biolog2$Subplot_Descriptions), FUN=mean)

#Plotting avg absorbance of each treatment across time
png("DGalacturonic_mean_L2.png", units="in", width=7, height=6, res=300)
ggplot(DGalacturonic_mean, aes(x = Group.1, y = x, group = Group.2, color = Group.2)) +
  geom_line(linewidth = 1) +
  geom_point(position=position_dodge(0.1)) + # "position = position_dodge()" is the code used to dodge overlapping points
  labs(y="D.Galacturonic Acid", x="Hour") +
  theme_bw() +
  theme(axis.title = element_text(size=17),
        axis.text = element_text(size=15))
dev.off()

```

Stacked bar plot showing Substrate average well color development (SAWCD) at hour 168
```{r}
# filter for hour 168
result168 <- result %>% filter(Hour == 168)

# calculate relative abundance
result168_relabun <- result168 %>%
  group_by(Subplot_Descriptions) %>%
  mutate(Relative_Abundance = avg_absorbence / sum(avg_absorbence))

# stacked bar plot
ggplot(result168_relabun, aes(x = Subplot_Descriptions, y = Relative_Abundance, fill = carbon_group)) +
        geom_bar(stat = "identity") +
        scale_fill_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E", "forestgreen", "darkblue")) +
        theme_minimal() +
        labs(x = "Treatment", y = "SAWCD")
```

PCAs
```{r}
biolog168 <- biolog %>% filter(Hour == 168)

pca_result <- rda(biolog168[,9:39], scale = TRUE)

# Extract PCA scores for the first two principal components
pca_scores <- as.data.frame(scores(pca_result, display = "sites"))

# Add sample meta data
pca_scores1 <- cbind(biolog168[,1:8], pca_scores[,1:2])

# Plot the PCA
ggplot(pca_scores1, aes(x = PC1, y = PC2, label = Subplot_Descriptions, color = Subplot_Descriptions)) +
  geom_point(size = 4) +
  geom_text(vjust = -1) +
  theme_minimal() +
  labs(x = "PC1",
       y = "PC2")
```

```{r}
# Aggregate the carbon source data by Subplot_Descriptions
biolog_agg <- biolog3 %>%
  group_by(Subplot_Descriptions) %>%
  summarise_if(is.numeric, mean, na.rm = TRUE)

# Run the PCA on the substrate data
pca_result1 <- prcomp(biolog_agg[, -c(1:5)], scale. = TRUE)  # Exclude the all columns except the subtrates

# Extract PCA scores (for plotting points)
pca_scores2 <- as.data.frame(pca_result1$x)
pca_scores2$Subplot_Descriptions <- biolog_agg$Subplot_Descriptions  # Add climate treatment info

# Extract PCA loadings (for plotting arrows)
pca_loadings <- as.data.frame(pca_result1$rotation)
pca_loadings$Substrate_Group <- rownames(pca_loadings)  # Add substrate group labels

# Plot the PCA results
ggplot(pca_scores2, aes(x = PC1, y = PC2, color = Subplot_Descriptions)) +
        geom_point(size = 5) +  # Points representing climate treatments
        geom_segment(data = pca_loadings, aes(x = 0, y = 0, xend = PC1, yend = PC2),
                     arrow = arrow(length = unit(0.3, "cm")), color = "black") +  # Arrows for substrate groups
        geom_text(data = pca_loadings, aes(x = PC1, y = PC2, label = Substrate_Group), 
                  vjust = 1, hjust = 1, color = "black", size = 3.5) +  # Labels for the arrows
        scale_color_manual(values = c("#a6bddb", "#2F4F4f", "#8B6914", "#9D422E", "forestgreen", "darkblue")) +
        labs(x = paste0("PC 1 (", round(summary(pca_result1)$importance[2, 1] * 100, 1), "%)"),
             y = paste0("PC 2 (", round(summary(pca_result1)$importance[2, 2] * 100, 1), "%)")) +
        theme_minimal() +
        theme(legend.position = "right")
```


