---
title: "Biolog Analysis Script"
author: "Adrian Noecker, Moriah Young"
date: "September 2023, updated January 2024"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
COLLABORATORS: Phoebe Zarnetske, Mark Hammond, Taylor Ulbrich (using code from Taylor)
DATA INPUT: L1 csv from the shared Google drive  
DATA OUTPUT: L2 csv and Code and Rmd are in the scripts folder in Github  
PROJECT: REX
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

#load packages
library(car)
library(tidyr)
library(dplyr)
library(tibble)
library(plyr)
library(readr)
library(tidyverse)
library(ggplot2)
library(lsmeans)
library(lme4)
library(tibble)
library(ggpubr)
library(lmerTest)
library(fitdistrplus)
library(sjPlot)
library(tidyverse)
library(car)
library(emmeans)
library(bbmle)
library(multcomp)
library(vegan)

# Set working directory from .Renviron
dir <- Sys.getenv("DATA_DIR")
list.files(dir)

## Read in data
# Adrian
biolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/biolog_clean_wide_format_L1.csv")) 
biolog2 <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/biolog_clean_L1.csv"))
diversity <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/biolog_hill_diversity_L2.csv"))

# Moriah's below (for some reason above wont work for me - not same file path)
biolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/biolog_clean_wide_format_L1.csv"))
biolog2 <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/biolog_clean_L1.csv"))
diversity <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L2/biolog_hill_diversity_L2.csv"))
```

# DATA ANALYSIS

Notes from Moriah:
- Model structure is a little more complex in REX - need to have random effects as such: (1|Rep/Footprint/Subplot). This is a nested plot design. For lack of better explanation, it goes from largest "scale" (rep) to smallest "scale" (subplot) This means that you'll want to retain the rep and footprint numbers in the dataframe you ultimately create below and with the diversity metrics.

#Looking at distribution of data

# this section was taken from Tayler's code - can probably delete
```{r}
###Starting with final reading (hour=168) 
biologHr168diversity <- diversity %>% filter(Hour == "168")

# test for normal residuals and equal variances (ANOVA assumptions)
# normal residuals?
ggqqplot(biologHr168diversity$H0)
hist(biologHr168diversity$H0) ##slightly skewed left

# Shapiro test - null hypothesis: residuals are normally distributed 
require(lme4)
require(lmerTest)
model = lm(H0 ~ Subplot_Descriptions, data = biologHr168diversity)
res = model$residuals
shapiro.test(res) 
#p = 0.002237 so not normally distributed?

# homogeneity of variances
# Many statistical tests (like a one-way ANOVA) assume that variances are equal across samples. Bartlett’s test can be 
# used to verify that assumption.
# H0: The variance among each group is equal.
# HA: At least one group has a variance that is not equal to the rest.

bartlett.test(H0 ~ Subplot_Descriptions, data = biologHr168diversity) 
# 0.6396 = we fail to reject the null hypothesis, so all groups have the same variance. This is what we want.

#ANOVA
Anova(model, type = "III")

#Response: Richness for hour 168
#                     Sum Sq Df F value    Pr(>F)    
#(Intercept)          3750.0  1 79.9290 2.009e-08 ***
#Subplot_Descriptions   53.0  3  0.3766    0.7709    
#Residuals             938.3 20   

# confidence intervals for every pairwise comparison
tukey <- TukeyHSD(aov(H0 ~ Subplot_Descriptions, data = biologHr168diversity))
tukey
plot(tukey)

#Subplot_Descriptions
#                            diff        lwr       upr     p adj
#drought-ambient        -0.500000 -11.568683 10.568683 0.9992524
#warmed-ambient         -3.833333 -14.902016  7.235349 0.7680882
#warmed_drought-ambient -1.000000 -12.068683 10.068683 0.9941408
#warmed-drought         -3.333333 -14.402016  7.735349 0.8334310
#warmed_drought-drought -0.500000 -11.568683 10.568683 0.9992524
#warmed_drought-warmed   2.833333  -8.235349 13.902016 0.8893834

```

Example script from other REX data on data analysis with steps for checking for assumption
https://github.com/kbs-lter/rex-analysis/blob/main/R/T7_warmx_plant_traits/Soca_plant_height_analyses_L2.R

# Data exploration
```{r}
biolog168 <- filter(biolog2, Hour == 168)
descdist(biolog168$Absorbence, discrete = FALSE)
hist(biolog168$Absorbence)
qqnorm(biolog168$Absorbence) # should be a straight diagonal line if normal
shapiro.test(biolog168$Absorbence) # p > 0.05 if normal

# need to do some sort of transformation

# sqrt transformation
biolog168$sqrt_Absorbence <- sqrt(biolog168$Absorbence)
descdist(biolog168$sqrt_Absorbence, discrete = FALSE)
hist(biolog168$sqrt_Absorbence)
qqnorm(biolog168$sqrt_Absorbence)
shapiro.test(biolog168$sqrt_Absorbence) # looks good

anova_result <- aov(carbon_group ~ Subplot_Descriptions, data = biolog168)
print(kruskal_test)
```

# LMER for final reads
```{r}
# without checking assumptions - need to do this. Just laying out the mixed effects model
m2 <- lmer(Shannon ~ Subplot_Descriptions + (1|Rep/Footprint_Location), data = biologshannon168, REML=F)
summary(m2) #looking at just hour 168 is singular
plot(m2, main = Shannon)

#levene test for homogeneity of variance
leveneTest(residuals(m2) ~ biologshannon168$Subplot_Descriptions) 

# Check for normal residuals
qqPlot(resid(m2), main = "ShannonAll")
hist(residuals(m2), main = "ShannonAll")
shapiro.test(resid(m2)) #p=0.02 - not normal but close

#anova
anova(m2)

#get rid of unneeded columns
meta2 <- meta1 %>% dplyr::select(Unique_ID, Rep, Footprint_Location)

#merge with meta for lmer
biologdiversity1 <- merge(x=biolog_diversity, y=meta2, by.x=c('Plot_ID'), 
      by.y=c('Unique_ID'))

# Shannon LMER with random effects
m3 <- lmer(Shannon ~ Subplot_Descriptions + Hour + (1|Rep/Footprint_Location), data = biologdiversity1, REML=F)
summary(m3)
plot(m3, main = Shannon)

#levene test for homogeneity of variance
leveneTest(residuals(m3) ~ biologdiversity1$Subplot_Descriptions) 

# Check for normal residuals
qqPlot(resid(m3), main = "Shannon")
hist(residuals(m3), main = "Shannon")
shapiro.test(resid(m3)) #p=0.3091

#anova
anova(m3)

# Richness LMER with random effects
m4 <- lmer(Richness ~ Subplot_Descriptions + Hour + (1|Rep/Footprint_Location), data = biologdiversity1, REML=F)
summary(m4)
plot(m4, main = Richness)

#levene test for homogeneity of variance
leveneTest(residuals(m4) ~ biologdiversity1$Subplot_Descriptions) 

# Check for normal residuals
qqPlot(resid(m4), main = "Richness")
hist(residuals(m4), main = "Richness")
shapiro.test(resid(m4)) #p=0.2442

#anova
anova(m4)

# AWCD LMER with random effects
m5 <- lmer(AWCD ~ Subplot_Descriptions + Hour + (1|Rep/Footprint_Location), data = biologdiversity1, REML=F)
summary(m5)
plot(m5, main = AWCD)

#levene test for homogeneity of variance
leveneTest(residuals(m5) ~ biologdiversity1$Subplot_Descriptions) 

# Check for normal residuals
qqPlot(resid(m5), main = "AWCD")
hist(residuals(m5), main = "AWCD")
shapiro.test(resid(m5)) #p=0.2366

#anova
anova(m5)

```

# this is an example:
# m3 <- lmer(Shannon ~ Subplot_Descriptions + hour + (1|Rep/Footprint_Location), data = biolog_diversity, REML=F)

## WORK IN PROGRESS!!
# Need to go back to when the matrices were made for each our and make sure the "subplot_descriptions" are included
## PCA with arrows
```{r}
# select only for hour 168
biolog.samples.Hr168 <- filter(biolog2, Hour == "168")

# make sample names the rownames so we keep them during ordination 
biolog.matrix.Hr168 <- column_to_rownames(biolog.samples.Hr168, "Plot_ID")


############ Multivariate analysis ##################
map<-biolog.matrix.Hr168[,1:5] #Converting  data to two matrices: Mapping and substrate (Comm)
comm<-biolog.matrix.Hr168[,6:35]
comm<-comm/rowSums(comm) #Relative abundance

# distance of rel. abund. matrix
dist<-vegdist(comm, method='bray', na.rm=TRUE)

#Principal component analysis
apcoa<-cmdscale(dist, k=2, eig=FALSE, add=FALSE, x.ret=FALSE)
plot<-ordiplot(apcoa, type="n")
vecpcoa<-envfit(apcoa,comm)
plot(vecpcoa)# add vectors
vecpcoa$vectors

# Set the colors by treatment 

points(plot, "sites", pch = 20, col = "green4", map$treatment=="SL_0") 
points(plot, "sites", pch = 20, col = "chartreuse",map$treatment=="SL_SL") 
points(plot, "sites", pch = 20, col = "blue1",map$treatment=="SL_BBS") 
points(plot, "sites", pch = 20, col = "darkmagenta",map$treatment=="SL_JG")
points(plot, "sites", pch = 20, col = "gold",map$treatment=="SL_S") 


plot + theme(legend.text = element_text(size = 16))

```

