---
title: "Biolog Analysis Script"
author: "Adrian Noecker, Moriah Young"
date: "September 2023, updated January 2024"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
COLLABORATORS: Phoebe Zarnetske, Mark Hammond, Taylor Ulbrich (using code from Taylor)
DATA INPUT: L1 csv from the shared Google drive  
DATA OUTPUT: L2 csv and Code and Rmd are in the scripts folder in Github  
PROJECT: REX
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

#load packages
library(car)
library(tidyr)
library(dplyr)
library(tibble)
library(plyr)
library(readr)
library(tidyverse)
library(ggplot2)
library(lsmeans)
library(lme4)
library(tibble)
library(ggpubr)
library(lmerTest)
library(fitdistrplus)
library(sjPlot)
library(tidyverse)
library(car)
library(emmeans)
library(bbmle)
library(multcomp)
library(vegan)

# Set working directory from .Renviron
dir <- Sys.getenv("DATA_DIR")
list.files(dir)

## Read in data
# Adrian
biolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/biolog_clean_wide_format_L1.csv")) 
biolog2 <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/biolog_clean_L1.csv"))
diversity <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/biolog_hill_diversity_L2.csv"))

# Moriah's below (for some reason above wont work for me - not same file path)
biolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/biolog_clean_wide_format_L1.csv"))
biolog2 <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/biolog_clean_L1.csv"))
diversity <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L2/biolog_hill_diversity_L2.csv"))
```

# DATA ANALYSIS

Notes from Moriah:
- Model structure is a little more complex in REX - need to have random effects as such: (1|Rep/Footprint/Subplot). This is a nested plot design. For lack of better explanation, it goes from largest "scale" (rep) to smallest "scale" (subplot) This means that you'll want to retain the rep and footprint numbers in the dataframe you ultimately create below and with the diversity metrics.

#Looking at distribution of data

# this section was taken from Tayler's code - can probably delete
```{r}
###Starting with final reading (hour=168) 
biologHr168diversity <- diversity %>% filter(Hour == "168")

# test for normal residuals and equal variances (ANOVA assumptions)
# normal residuals?
ggqqplot(biologHr168diversity$H0)
hist(biologHr168diversity$H0) ##slightly skewed left

# Shapiro test - null hypothesis: residuals are normally distributed 
require(lme4)
require(lmerTest)
model = lm(H0 ~ Subplot_Descriptions, data = biologHr168diversity)
res = model$residuals
shapiro.test(res) 
#p = 0.002237 so not normally distributed?

# homogeneity of variances
# Many statistical tests (like a one-way ANOVA) assume that variances are equal across samples. Bartlett’s test can be 
# used to verify that assumption.
# H0: The variance among each group is equal.
# HA: At least one group has a variance that is not equal to the rest.

bartlett.test(H0 ~ Subplot_Descriptions, data = biologHr168diversity) 
# 0.6396 = we fail to reject the null hypothesis, so all groups have the same variance. This is what we want.

#ANOVA
Anova(model, type = "III")

#Response: Richness for hour 168
#                     Sum Sq Df F value    Pr(>F)    
#(Intercept)          3750.0  1 79.9290 2.009e-08 ***
#Subplot_Descriptions   53.0  3  0.3766    0.7709    
#Residuals             938.3 20   

# confidence intervals for every pairwise comparison
tukey <- TukeyHSD(aov(H0 ~ Subplot_Descriptions, data = biologHr168diversity))
tukey
plot(tukey)

#Subplot_Descriptions
#                            diff        lwr       upr     p adj
#drought-ambient        -0.500000 -11.568683 10.568683 0.9992524
#warmed-ambient         -3.833333 -14.902016  7.235349 0.7680882
#warmed_drought-ambient -1.000000 -12.068683 10.068683 0.9941408
#warmed-drought         -3.333333 -14.402016  7.735349 0.8334310
#warmed_drought-drought -0.500000 -11.568683 10.568683 0.9992524
#warmed_drought-warmed   2.833333  -8.235349 13.902016 0.8893834

```

Help from stack exchange - cross validated
https://stats.stackexchange.com/questions/659832/how-to-analyze-a-dataset-with-a-response-variable-thats-continuous-and-has-a-lo
```{r}
biolog168 <- filter(biolog2, Hour == 168)
library(gamlss)

# create function that normalizes data
normalize <- function(x) (x - min(x)) / (max(x) - min(x))
# run data through the function
biolog168$Absorbence <- normalize(biolog168$Absorbence)

# set up the model
mod <- gamlss(
  Absorbence ~ Subplot_Descriptions + # fixed effect
    re(random = ~ 1 | Rep / Footprint_Location), # random effects
  family = BEINF(), # zero- and one-inflated distribution
  data = biolog168 
)

plot(mod)

summary(mod)

# calculate effect size
# Coefficients
intercept <- -0.83045
drought_coeff <- -0.28444

# Calculate mu for control
mu_control <- exp(intercept) / (1 + exp(intercept))

# Calculate mu for drought
mu_drought <- exp(intercept + drought_coeff) / (1 + exp(intercept + drought_coeff))

# Effect size
effect_size <- mu_control - mu_drought
effect_size #0.05658956
# This means the drought treatment reduced mean Absorbance by approximately 5.7% compared to the control.

ggplot(biolog168, aes(x = Subplot_Descriptions, y = Absorbence)) +
        geom_violin()
```
We analyzed the effects of subplot treatments on Absorbance using a generalized additive model for location, scale, and shape (GAMLSS) with a Beta Inflated (BEINF) distribution. This distribution was appropriate given the response variable's bounded nature (0–1) and potential for inflation at 0 or 1. Fixed effects included subplot treatment (Subplot_Descriptions), and random effects accounted for nested structure within replicates (Rep / Footprint_Location).

The mean Absorbance (μ) was significantly affected by treatment (p=0.0337for drought). Compared to the control, drought treatments reduced Absorbance (Estimate = -0.284, p=0.0337). In contrast, warmed and warmed_drought treatments did not significantly differ from the control (Estimate = -0.039, p=0.7796; Estimate = 0.180, p=0.1821, respectively).

The dispersion (σ) parameter, which captures variability in Absorbance around the mean, was significant overall (Intercept Estimate = 0.504, p<0.001), indicating moderate variability under the reference condition. However, treatment effects on dispersion were not modeled directly.

Probabilities of zero-inflation (ν) and one-inflation (τ) were both low under the reference condition, as indicated by negative intercept estimates (ν: Estimate = -1.170, p<0.001; τ: Estimate = -6.340, p<0.001).

The inclusion of random effects accounted for variability among replicates and within footprints, improving model fit and addressing non-independence in the data. The model explained the data well, as evidenced by a Global Deviance of 191.89 and an AIC of 223.20.

Interpretation
These results suggest that drought treatments significantly reduced mean Absorbance compared to control subplots, while other treatments (warmed and warmed_drought) had no detectable effect. This indicates a specific sensitivity of the system to drought conditions. Variability in Absorbance was moderate but consistent across treatments, and the probabilities of 0 or 1-inflation in Absorbance were negligible under all conditions.

```{r}
mod2 <- gamlss(
  Absorbence ~ Subplot_Descriptions * carbon_group + # fixed effect
    re(random = ~ 1 | Rep / Footprint_Location), # random effects
  family = BEINF(), # zero- and one-inflated distribution
  data = biolog168 
)
plot(mod2)
summary(mod2)
```

diversity

richness - H0
```{r}
descdist(diversity$H0, discrete = FALSE)
hist(diversity$H0)
qqnorm(diversity$H0) # should be a straight diagonal line if normal
shapiro.test(diversity$H0) # p > 0.05 if normal

# test for normal residuals and equal variances (ANOVA assumptions)
# normal residuals?
ggqqplot(diversity$H0)
hist(diversity$H0) ##slightly skewed left

# Shapiro test - null hypothesis: residuals are normally distributed 
require(lme4)
require(lmerTest)
model = lm(H0 ~ Subplot_Descriptions, data = diversity)
res = model$residuals
shapiro.test(res) 
# not normally distributed

bartlett.test(H0 ~ Subplot_Descriptions, data = diversity) 

# try some transformations
# sqrt transformation
diversity$sqrt_H0 <- sqrt(diversity$H0)
descdist(diversity$sqrt_H0, discrete = FALSE)
hist(diversity$sqrt_H0)
qqnorm(diversity$sqrt_H0)
shapiro.test(diversity$sqrt_H0) 

# log transformation
diversity$log_H0 <- log(diversity$H0)
descdist(diversity$log_H0, discrete = FALSE)
hist(diversity$log_H0)
qqnorm(diversity$log_H0)
shapiro.test(diversity$log_H0) 

# no transformations look the best
```

shannon diversity - H1
```{r}
descdist(diversity$H1, discrete = FALSE)
hist(diversity$H1)
qqnorm(diversity$H1) # should be a straight diagonal line if normal
shapiro.test(diversity$H1) # p > 0.05 if normal

# log transformation
diversity$log_H1 <- log(diversity$H1)
descdist(diversity$log_H1, discrete = FALSE)
hist(diversity$log_H1)
qqnorm(diversity$log_H1)
shapiro.test(diversity$log_H1) 
```







# can ignore below

non parametric test but no random effects
```{r}
kruskal.test(Absorbence ~ Subplot_Descriptions, data = biolog168)

library(dunn.test)

# Perform Dunn's Test for a post-hoc analysis
dunn_result <- dunn.test(biolog168$Absorbence, biolog168$Subplot_Descriptions, method = "bonferroni")
```

```{r}
# Rank-transform the response variable
biolog168$RankAbsorbence <- rank(biolog168$Absorbence)

ggqqplot(biolog168$RankAbsorbence)
descdist(biolog168$RankAbsorbencee, discrete = FALSE)
hist(biolog168$RankAbsorbence)
qqnorm(biolog168$RankAbsorbence) # should be a straight diagonal line if normal
shapiro.test(biolog168$RankAbsorbence) # p > 0.05 if normal

# Fit a mixed-effects model with the rank-transformed response
lmm <- lmer(RankAbsorbence ~ Subplot_Descriptions + (1|Rep/Footprint_Location), data = biolog168)

# Summary of the model
anova(lmm)
summary(lmm)
```

Example script from other REX data on data analysis with steps for checking for assumption
https://github.com/kbs-lter/rex-analysis/blob/main/R/T7_warmx_plant_traits/Soca_plant_height_analyses_L2.R

# Data exploration
```{r}
biolog168 <- filter(biolog2, Hour == 168)
descdist(biolog168$Absorbence, discrete = FALSE)
hist(biolog168$Absorbence)
qqnorm(biolog168$Absorbence) # should be a straight diagonal line if normal
shapiro.test(biolog168$Absorbence) # p > 0.05 if normal

# need to do some sort of transformation

# sqrt transformation
biolog168$sqrt_Absorbence <- sqrt(biolog168$Absorbence)
descdist(biolog168$sqrt_Absorbence, discrete = FALSE)
hist(biolog168$sqrt_Absorbence)
qqnorm(biolog168$sqrt_Absorbence)
shapiro.test(biolog168$sqrt_Absorbence) 

# log transformation
biolog168$log_Absorbence <- log(biolog168$Absorbence + 1)
descdist(biolog168$log_Absorbence, discrete = FALSE)
hist(biolog168$log_Absorbence)
qqnorm(biolog168$log_Absorbence)
shapiro.test(biolog168$log_Absorbence) 

biolog168$inverse_Absorbence <- 1 / (biolog168$Absorbence + 1e-6)
descdist(biolog168$inverse_Absorbence, discrete = FALSE)
hist(biolog168$inverse_Absorbence)
qqnorm(biolog168$inverse_Absorbence)
shapiro.test(biolog168$inverse_Absorbence) 

```

# LMER for final reads
```{r}
# without checking assumptions - need to do this. Just laying out the mixed effects model
m2 <- lmer(Shannon ~ Subplot_Descriptions + (1|Rep/Footprint_Location), data = biologshannon168, REML=F)
summary(m2) #looking at just hour 168 is singular
plot(m2, main = Shannon)

#levene test for homogeneity of variance
leveneTest(residuals(m2) ~ biologshannon168$Subplot_Descriptions) 

# Check for normal residuals
qqPlot(resid(m2), main = "ShannonAll")
hist(residuals(m2), main = "ShannonAll")
shapiro.test(resid(m2)) #p=0.02 - not normal but close

#anova
anova(m2)

#get rid of unneeded columns
meta2 <- meta1 %>% dplyr::select(Unique_ID, Rep, Footprint_Location)

#merge with meta for lmer
biologdiversity1 <- merge(x=biolog_diversity, y=meta2, by.x=c('Plot_ID'), 
      by.y=c('Unique_ID'))

# Shannon LMER with random effects
m3 <- lmer(Shannon ~ Subplot_Descriptions + Hour + (1|Rep/Footprint_Location), data = biologdiversity1, REML=F)
summary(m3)
plot(m3, main = Shannon)

#levene test for homogeneity of variance
leveneTest(residuals(m3) ~ biologdiversity1$Subplot_Descriptions) 

# Check for normal residuals
qqPlot(resid(m3), main = "Shannon")
hist(residuals(m3), main = "Shannon")
shapiro.test(resid(m3)) #p=0.3091

#anova
anova(m3)

# Richness LMER with random effects
m4 <- lmer(Richness ~ Subplot_Descriptions + Hour + (1|Rep/Footprint_Location), data = biologdiversity1, REML=F)
summary(m4)
plot(m4, main = Richness)

#levene test for homogeneity of variance
leveneTest(residuals(m4) ~ biologdiversity1$Subplot_Descriptions) 

# Check for normal residuals
qqPlot(resid(m4), main = "Richness")
hist(residuals(m4), main = "Richness")
shapiro.test(resid(m4)) #p=0.2442

#anova
anova(m4)

# AWCD LMER with random effects
m5 <- lmer(AWCD ~ Subplot_Descriptions + Hour + (1|Rep/Footprint_Location), data = biologdiversity1, REML=F)
summary(m5)
plot(m5, main = AWCD)

#levene test for homogeneity of variance
leveneTest(residuals(m5) ~ biologdiversity1$Subplot_Descriptions) 

# Check for normal residuals
qqPlot(resid(m5), main = "AWCD")
hist(residuals(m5), main = "AWCD")
shapiro.test(resid(m5)) #p=0.2366

#anova
anova(m5)

```

