---
title: "Biolog Diversity Calculations Script"
author: "Adrian Noecker, Moriah Young"
date: "September 2023, updated February 2024"
output: pdf_document
editor_options: 
  chunk_output_type: console
---
COLLABORATORS: Phoebe Zarnetske, Mark Hammond, Taylor Ulbrich (using code from Taylor)
DATA INPUT: L1 csv from the shared Google drive  
DATA OUTPUT: L2 csv and Code and Rmd are in the scripts folder in Github  
PROJECT: REX
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Clear all existing data
rm(list=ls())

#load packages
library(car)
library(tidyr)
library(dplyr)
library(tibble)
library(plyr)
library(readr)
library(tidyverse)
library(ggplot2)
library(tibble)
library(ggpubr)
library(sjPlot)
library(tidyverse)
library(bbmle)
library(multcomp)
library(vegan)

# Set working directory from .Renviron
dir <- Sys.getenv("DATA_DIR")
list.files(dir)
setwd(dir)

## Read in data
biolognowater <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/Biolog_data_NOWater_L1.csv")) # data in wide format, water not included
standardbiolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates/T7_warmx/L1/Biolog_data_L1.csv")) # data in long format, water included

# Moriah's below (for some reason above wont work for me - not same file path)
biolognowater <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/Biolog_data_NOWater_L1.csv"))
standardbiolog <- read.csv(file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L1/Biolog_data_L1.csv"))
```

** MOVE THIS TO CLEAN SCRIPT **
# reformatting and checking absorbances
# Note: the "biolog" dataframe is in wide format, and is the format that the data is output from the plate reader. The "standard" data frame is in narrow format.
```{r}
#removing water from standard biolog
standardnowater <- standardbiolog %>% filter(!Well_C_Source %in% c("Water"))

#changing row name and reformatting, combine hour and ID to one column
biologdf <- biolognowater %>%
  unite(SampleID_Hour, Biolog_column_ID, Hour, sep ="_", remove = FALSE)

standarddf <- standardnowater %>%
  unite(SampleID_Hour, Biolog_column_ID, Hour, sep ="_", remove = FALSE)

#renaming new column
biologdf$rowname <- biologdf$SampleID_Hour

# Note from Moriah: is there a reason why the first column is converted to the row names??
#changing column to row
biologmatrix <- column_to_rownames(biologdf, "rowname")

#Check control absorbances (all should == 0 if no contamination)
biologcontrol <- biologmatrix %>% filter(Sample_ID=="BIOLOG CONTROL")
####Tween 40, Tween 80, D Xylose, X4 Hydroxy Benzoic Acid, and X2 Hydroxy Benzoic Acid larger than 0.01 in a few controls, most hour 0 in plates 5, 7, 9, and 10 so maybe just the initial color of well and not contamination??

standardcontrol <- standarddf %>% filter(Sample_ID=="BIOLOG CONTROL")

#exclude controls for analysis
biologsamples <- biologmatrix %>% filter(!Sample_ID %in% c("BIOLOG CONTROL"))

standardsamples <- standarddf %>% filter(!Sample_ID %in% c("BIOLOG CONTROL"))

#import meta data to get treatment type
meta <- read.csv(file.path(dir, "/REX_warmx_metadata.csv"))

#get rid of unneeded columns
meta1 <- meta %>% dplyr::select(Subplot_Descriptions, Unique_ID, Rep, Footprint_Location)

# Note from Moriah: when we merge these two together, we lose the matrix formatting - may need to run "column_to_rownames" function again on new dataframe to make SampleID_Hour row names
#combine meta and sample df
biologsamples1 <- merge(x=biologsamples, y=meta1, by.x=c('Plot_ID'), 
      by.y=c('Unique_ID'))

standardsamples1 <- merge(x=standardsamples, y=meta1, by.x=c('Plot_ID'), 
      by.y=c('Unique_ID'))

#change column order
biologsamples2 <- biologsamples1 %>% relocate(Subplot_Descriptions, .after = Biolog_column_ID) %>% 
        relocate(Rep, .after = Plot_ID) %>% 
        relocate(Footprint_Location, .after = Rep)

# upload new data frame to L2 folder in warmx 
write.csv(biologsamples2, file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L2/biolog_hill_diversity_L2.csv"), row.names = FALSE)

```

Rarefaction
These aren't counts though, they are absorbency levels so I don't think rarefaction will work on this unless we convert
to presence absence?
**come back to this
```{r}
# calculate total number of trees per sample
biolog.rowsums <- apply(biologsamples2[,-c(1:8)],1,sum)
# whats the value of smallest row sum?
min(biolog.rowsums) # 2.062 - I'll use this when I rarefy

###Individual based rarefaction
# Using the rarefaction formula to find expected species richness
rarecurve(biologsamples2[,-c(1:8)])
# You can choose a minimum sample size to rarefy all samples to using the command below
rarefy(biologsamples2[,-c(1:8)],sample=2)
```

#grouping by hour
```{r}
# split data by Hour so we can calculate shannon diversity, richness, and AWCD by each Hour 
biologHr0 <- filter(biologsamples2, Hour == 0)
biologHr24 <- filter(biologsamples2, Hour == 24)
biologHr48 <- filter(biologsamples2, Hour == 48)
biologHr72 <- filter(biologsamples2, Hour == 72)
biologHr96 <- filter(biologsamples2, Hour == 96)
biologHr120 <- filter(biologsamples2, Hour == 120)
biologHr144 <- filter(biologsamples2, Hour == 144)
biologHr168 <- filter(biologsamples2, Hour == 168)
```

# Calculate Hill numbers and avg AWCD
# might need to come back to this - do we need to get averages of hill numbers since they are slightly different every time?
```{r}
#biolog.hill <- renyi(biologHr168[,-c(1:8)], scales=c(0,1,2,Inf), hill=TRUE)
## calculate Average Well Color Development (mean color change across all 31 substrates)
#AWCD <- apply(biologHr168[,-c(1:8)],1,mean)
#biologHr168.hill <- cbind(biologHr168[1:8], biolog.hill, AWCD)

# create a function to calculate hill numbers and average AWCD across all the hours
renyi_function <- function(data) {
  biolog.hill <- renyi(data[,-c(1:8)], scales = c(0,1,2,Inf), hill = TRUE)
  avg_AWCD <- apply(data[,-c(1:8)], 1, mean)
  biolog.hill.AWCD <- cbind(data[1:8], biolog.hill, avg_AWCD)
}

biologHr168.hill <- renyi_function(biologHr168)
biologHr144.hill <- renyi_function(biologHr144)
biologHr120.hill <- renyi_function(biologHr120)
biologHr96.hill <- renyi_function(biologHr96)
biologHr72.hill <- renyi_function(biologHr72)
biologHr48.hill <- renyi_function(biologHr48)
biologHr24.hill <- renyi_function(biologHr24)
biologHr0.hill <- renyi_function(biologHr0)

# merge these dataframes that were created above together into one dataframe
# put all data frames into list
biolog_list <- list(biologHr168.hill, biologHr144.hill, biologHr120.hill, biologHr96.hill,
                              biologHr72.hill, biologHr48.hill, biologHr24.hill, biologHr0.hill)

# merge all data frames in list and create new dataframe
biolog_hill_diversity <- biolog_list %>% reduce(full_join)

# rename column names for the hill numbers
colnames(biolog_hill_diversity)[9] ="H0"
colnames(biolog_hill_diversity)[10] ="H1"
colnames(biolog_hill_diversity)[11] ="H2"
colnames(biolog_hill_diversity)[12] ="HInf"

# upload new data frame to L2 folder in warmx 
write.csv(biolog_hill_diversity, file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L2/biolog_hill_diversity_L2.csv"), row.names = FALSE)
```





# get rid of code below?
Diversity Calculations separated by hour
```{r}
###Hour 168
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness <- rowSums(biologHr168[,-c(1:8)]>0)

# Second, calculate shannon diversity 
Shannon <- diversity(biologHr168[,-c(1:8)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD <- apply(biologHr168[,-c(1:8)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr168diversity <- cbind(biologHr168[1:8], Richness, Shannon, AWCD)

###Hour 144
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness <- rowSums(biologHr144[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon <- diversity(biologHr144[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD <- apply(biologHr144[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr144diversity <- cbind(biologHr144[1:6], Richness, Shannon, AWCD)

###Hour 120
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness <- rowSums(biologHr120[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon <- diversity(biologHr120[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD <- apply(biologHr120[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr120diversity <- cbind(biologHr120[1:6], Richness, Shannon, AWCD)

###Hour 96
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness <- rowSums(biologHr96[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon <- diversity(biologHr96[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD <- apply(biologHr96[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr96diversity <- cbind(biologHr96[1:6], Richness, Shannon, AWCD)

###Hour 72
# First, richness of subsrates used; add all indices as a column to the metadata file 
Richness <- rowSums(biologHr72[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon <- diversity(biologHr72[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD <- apply(biologHr72[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr72diversity <- cbind(biologHr72[1:6], Richness, Shannon, AWCD)

###Hour 48
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness <- rowSums(biologHr48[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon <- diversity(biologHr48[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD <- apply(biologHr48[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr48diversity <- cbind(biologHr48[1:6], Richness, Shannon, AWCD)

###Hour 24
# First, richness of substrates used; add all indices as a column to the metadata file 
Richness <- rowSums(biologHr24[,-c(1:6)]>0)

# Second, calculate shannon diversity 
Shannon <- diversity(biologHr24[,-c(1:6)], index = "shannon")

# Third, calculate Average Well Color Development (mean color change across all 31 substrates)
AWCD2 <- apply(biologHr24[,-c(1:6)],1,mean)

# merge diversity data with SampleID_Hour for 168hr
biologHr24diversity <- cbind(biologHr24[1:6], Richness, Shannon, AWCD)

# merge these dataframes that were created above together into one dataframe
# put all data frames into list
biolog_list <- list(biologHr24diversity, biologHr48diversity, biologHr72diversity, biologHr96diversity,
                              biologHr120diversity, biologHr144diversity, biologHr168diversity)

# merge all data frames in list and create new dataframe
biolog_diversity <- biolog_list %>% reduce(full_join)

# upload new data frame to L2 folder in warmx 
#write.csv(biolog_diversity, file.path(dir, "/microbes/Biolog EcoPlates (1)/T7_warmx/L2/biolog_diversity_L2.csv"), row.names = FALSE)
```

